### 2005-05-09 docs/code_list.tex

\section{Simple Script}
Below is the code for the script used to run the SimpleScalar simulations.

\lgrindfile{code/simple_code}
\label{Simple Script}

\section{Sample Simulation Run}
\label{Simple List}
The instructions listed below form a sample simulation run of SimpleScalar. Each
of the commands is run in turn and its results stored by \cc{simple}.

\lgrindfile{code/sample_sim_list}

\section{Sample Simulation}
The text below is an output of a simgle run of \cc{sim-bpred}.
\label{Sample Simulation}

\lgrindfile{code/sample_output}

\section{Convert Script}
The following is the source code to the script used to convert data from the
stored format of the \cc{simple} script to that required by gnuplot.
\label{Convert Script}

\lgrindfile{code/convert}

\section{Sample Gnuplot Script}
\label{Sample Gnuplot Script}
The following script is used to make plot results. The image made by this file
is in Figure \vref{All cycles}.

\lgrindfile{code/cycles_script}

\section{Sort listings}
Included below is a code listing of the sorts used in this project. Note that
the code is not exact. It is edited for brevity and clarity, and may not even
compile. In addition, loops are rolled, edge-cases are removed and variable
declarations are moved or removed. The actual code used is contained on the CD
accompanying this project.
\label{Sort Listings}

Base heapsort is listed in Figure \ref{Simple Base Heapsort}.\\
\textbf{Cache Heapsort:}
\lgrindfile{code/cache_heapsort_code}

\textbf{Base Mergesort:}
\lgrindfile{code/base_mergesort_code}

Tiled Mergesort is very similar to double-tiled mergesort, and so is not
included. Merge and presort functions are also not included, as they are very
similar to those of base mergesort.\\
\textbf{Double Tiled Mergesort:}
\lgrindfile{code/double_tiled_mergesort_code}

\textbf{Double Multi-Mergesort:}
\lgrindfile{code/double_multi_mergesort_code}

\textbf{Base Quicksort:}
\lgrindfile{code/base_quicksort_code}

\textbf{Cache Quicksort:}
\lgrindfile{code/cache_quicksort_code}

\textbf{Multi-Quicksort:}
\lgrindfile{code/multi_quicksort_code}

\textbf{Sequential Multi-Quicksort:}
\lgrindfile{code/multi_quicksort_seq_code}

\textbf{Radixsort:}
\lgrindfile{code/radixsort_code}