### 2005-05-26 docs/demo.tex
make way for the new and improved fs

\documentclass[a4,landscape]{seminar}

\usepackage{semcolor}
\usepackage{graphicx}
\usepackage{color}
\usepackage{pst-all}

\definecolor{borderred}{rgb}{0.27,0.47,0.68}
\definecolor{darkblue}{rgb}{0.27,0.18,0.55}
\newenvironment{stitle}{\LARGE\color{black}}{}

\slideframe[\psset{fillstyle=gradient,gradmidpoint=0.25,gradbegin=white,gradend=gray,linecolor=white}]{scplain} %to show

%\slideframe{none} %to print

\renewcommand{\slidetopmargin}{-.2in}
\addtolength{\slidewidth}{2.4in}
\addtolength{\slideheight}{1.8in}

\newcommand{\plot}[2]{\begin{figure}\centering\includegraphics*[scale=#1]{plots/#2}\end{figure}}
\newcommand{\plotdouble}[3]{\begin{figure}\centering\includegraphics*[scale=#1]{plots/#2}\hspace{5pt}\includegraphics*[scale=#1]{plots/#3}\end{figure}}

\begin{document}
\centering
\sffamily
\input slide1.tex
\input slide2.tex
\input flowchart.tex
\input slide4.tex
\input slide5.tex
\input slide6.tex
\input slide7.tex
\input slide8.tex
\input slide3.tex

\end{document}