### 2005-05-26 paper/docs/bugs.tex

\section{Bugs}
This section lists some of the bugs that exist in the remaining code. In a
project of limited timescale, it is not possible to remove every possible bug
from the code. These bugs are listed here in order that it is possible to
continue on this work with minimal difficutly.

\section{Results}
\label{bug3}
To reduce the costs of creating the array of random numbers, this time was
measured and subtracted from the results. While this works as expected for
instruction count, the effects on cache misses is less clear. Subtracting level
2 cache misses incurred during filling removes compulsory misses from the
equation. For shorter lists, the compulsory misses are completely removed.
For longer lists, those greater than the size of the cache, compulsory misses
come back, since the keys at the start of the array are flushed from the cache
by the time they begin to be sorted. LaMarca's work centers around
reducing conflict misses, however, so this is not necessarily a problem, just
something to be aware of.