### 2005-05-26 paper/docs/code/sample_sim_list.tex

% Remember to use the lgrind style

\Head{}
\File{sample\_sim\_list}{2004}{4}{28}{3:48}{2185}
\L{\LB{\V{sslittle}\-\V{na}\-\V{sstrix}\-\V{gcc}_\-\V{Wall}_\-\V{O3}_\-\V{funroll}\-\V{loops}_\-\V{finline}\-\V{functions}_}}
\L{\LB{}\Tab{8}{\-\V{DRANDOM\_SIZE}=\N{4096}_\-\V{DSEED}=\N{0}_\-\V{DSORT}=\V{base\_heapsort}_}}
\L{\LB{}\Tab{8}{\V{base\_heapsort}.\V{c}_\V{simpleMain}.\V{c}_\-\V{o}_\V{base\_heapsort}}}
\L{\LB{\V{sim}\-\V{cache}_\-\V{cache}:\V{dl1}_\V{dl1}:\N{256}:\N{32}:\N{1}:\V{l}_\-\V{cache}:\V{dl2}_\V{dl2}:\N{131072}:\N{32}:\N{1}:\V{l}_}}
\L{\LB{}\Tab{24}{\-\V{cache}:\V{il1}_\V{il1}:\N{256}:\N{32}:\N{1}:\V{l}_\-\V{cache}:\V{il2}_\V{dl2}_\V{base\_heapsort}}}
\L{\LB{\V{sim}\-\V{cache}_\-\V{cache}:\V{dl1}_\V{dl1}:\N{256}:\N{32}:\N{1}:\V{l}_\-\V{cache}:\V{dl2}_\V{dl2}:\N{65536}:\N{32}:\N{1}:\V{l}_}}
\L{\LB{}\Tab{24}{\-\V{cache}:\V{il1}_\V{il1}:\N{256}:\N{32}:\N{1}:\V{l}_\-\V{cache}:\V{il2}_\V{dl2}_\V{base\_heapsort}}}
\L{\LB{\V{sim}\-\V{cache}_\-\V{cache}:\V{dl1}_\V{dl1}:\N{1}:\N{32}:\N{256}:\V{l}_\-\V{cache}:\V{dl2}_\V{dl2}:\N{1}:\N{32}:\N{131072}:\V{l}_}}
\L{\LB{}\Tab{24}{\-\V{cache}:\V{il1}_\V{il1}:\N{1}:\N{32}:\N{256}:\V{l}_\-\V{cache}:\V{il2}_\V{dl2}_\V{base\_heapsort}}}
\L{\LB{\V{sim}\-\V{cache}_\-\V{cache}:\V{dl1}_\V{dl1}:\N{1}:\N{32}:\N{256}:\V{l}_\-\V{cache}:\V{dl2}_\V{dl2}:\N{1}:\N{32}:\N{65536}:\V{l}_}}
\L{\LB{}\Tab{24}{\-\V{cache}:\V{il1}_\V{il1}:\N{1}:\N{32}:\N{256}:\V{l}_\-\V{cache}:\V{il2}_\V{dl2}_\V{base\_heapsort}}}
\L{\LB{\V{sim}\-\V{bpred}_\-\V{bpred}_\V{bimod}_\-\V{bpred}:\V{bimod}_\N{512}_\V{base\_heapsort}}}
\L{\LB{\V{sim}\-\V{bpred}_\-\V{bpred}_\V{bimod}_\-\V{bpred}:\V{bimod}_\N{1024}_\V{base\_heapsort}}}
\L{\LB{\V{sim}\-\V{bpred}_\-\V{bpred}_\V{bimod}_\-\V{bpred}:\V{bimod}_\N{2048}_\V{base\_heapsort}}}
\L{\LB{\V{sim}\-\V{bpred}_\-\V{bpred}_\N{2$\ell$ev}_\-\V{bpred}:\N{2$\ell$ev}_\N{1}_\N{512}_\N{10}_\N{1}_\V{base\_heapsort}}}
\L{\LB{\V{sim}\-\V{bpred}_\-\V{bpred}_\N{2$\ell$ev}_\-\V{bpred}:\N{2$\ell$ev}_\N{1}_\N{1024}_\N{10}_\N{1}_\V{base\_heapsort}}}
\L{\LB{\V{sim}\-\V{bpred}_\-\V{bpred}_\N{2$\ell$ev}_\-\V{bpred}:\N{2$\ell$ev}_\N{1}_\N{2048}_\N{10}_\N{1}_\V{base\_heapsort}}}
\L{\LB{\V{sim}\-\V{bpred}_\-\V{bpred}_\N{2$\ell$ev}_\-\V{bpred}:\N{2$\ell$ev}_\N{1}_\N{4096}_\N{10}_\N{1}_\V{base\_heapsort}}}
\L{\LB{}}
\L{\LB{}\Tab{56}{.}}
\L{\LB{}\Tab{56}{.}}
\L{\LB{}\Tab{56}{.}}
\L{\LB{}\Tab{56}{.}}
\L{\LB{}\Tab{56}{.}}
\L{\LB{}\Tab{56}{.}}
\L{\LB{}}
\L{\LB{\V{sslittle}\-\V{na}\-\V{sstrix}\-\V{gcc}_\-\V{Wall}_\-\V{O3}_\-\V{funroll}\-\V{loops}_\-\V{finline}\-\V{functions}_}}
\L{\LB{}\Tab{8}{\-\V{DRANDOM\_SIZE}=\N{4194304}_\-\V{DSEED}=\N{0}_\-\V{DSORT}=\V{base\_heapsort}_}}
\L{\LB{}\Tab{8}{\V{base\_heapsort}.\V{c}_\V{simpleMain}.\V{c}_\-\V{o}_\V{base\_heapsort}}}
\L{\LB{\V{sim}\-\V{cache}_\-\V{cache}:\V{dl1}_\V{dl1}:\N{256}:\N{32}:\N{1}:\V{l}_\-\V{cache}:\V{dl2}_\V{dl2}:\N{131072}:\N{32}:\N{1}:\V{l}_}}
\L{\LB{}\Tab{24}{\-\V{cache}:\V{il1}_\V{il1}:\N{256}:\N{32}:\N{1}:\V{l}_\-\V{cache}:\V{il2}_\V{dl2}_\V{base\_heapsort}}}
\L{\LB{\V{sim}\-\V{cache}_\-\V{cache}:\V{dl1}_\V{dl1}:\N{256}:\N{32}:\N{1}:\V{l}_\-\V{cache}:\V{dl2}_\V{dl2}:\N{65536}:\N{32}:\N{1}:\V{l}_}}
\L{\LB{}\Tab{24}{\-\V{cache}:\V{il1}_\V{il1}:\N{256}:\N{32}:\N{1}:\V{l}_\-\V{cache}:\V{il2}_\V{dl2}_\V{base\_heapsort}}}
\L{\LB{\V{sim}\-\V{cache}_\-\V{cache}:\V{dl1}_\V{dl1}:\N{1}:\N{32}:\N{256}:\V{l}_\-\V{cache}:\V{dl2}_\V{dl2}:\N{1}:\N{32}:\N{131072}:\V{l}_}}
\L{\LB{}\Tab{24}{\-\V{cache}:\V{il1}_\V{il1}:\N{1}:\N{32}:\N{256}:\V{l}_\-\V{cache}:\V{il2}_\V{dl2}_\V{base\_heapsort}}}
\L{\LB{\V{sim}\-\V{cache}_\-\V{cache}:\V{dl1}_\V{dl1}:\N{1}:\N{32}:\N{256}:\V{l}_\-\V{cache}:\V{dl2}_\V{dl2}:\N{1}:\N{32}:\N{65536}:\V{l}_}}
\L{\LB{}\Tab{24}{\-\V{cache}:\V{il1}_\V{il1}:\N{1}:\N{32}:\N{256}:\V{l}_\-\V{cache}:\V{il2}_\V{dl2}_\V{base\_heapsort}}}
\L{\LB{\V{sim}\-\V{bpred}_\-\V{bpred}_\V{bimod}_\-\V{bpred}:\V{bimod}_\N{512}_\V{base\_heapsort}}}
\L{\LB{\V{sim}\-\V{bpred}_\-\V{bpred}_\V{bimod}_\-\V{bpred}:\V{bimod}_\N{1024}_\V{base\_heapsort}}}
\L{\LB{\V{sim}\-\V{bpred}_\-\V{bpred}_\V{bimod}_\-\V{bpred}:\V{bimod}_\N{2048}_\V{base\_heapsort}}}
\L{\LB{\V{sim}\-\V{bpred}_\-\V{bpred}_\N{2$\ell$ev}_\-\V{bpred}:\N{2$\ell$ev}_\N{1}_\N{512}_\N{10}_\N{1}_\V{base\_heapsort}}}
\L{\LB{\V{sim}\-\V{bpred}_\-\V{bpred}_\N{2$\ell$ev}_\-\V{bpred}:\N{2$\ell$ev}_\N{1}_\N{1024}_\N{10}_\N{1}_\V{base\_heapsort}}}
\L{\LB{\V{sim}\-\V{bpred}_\-\V{bpred}_\N{2$\ell$ev}_\-\V{bpred}:\N{2$\ell$ev}_\N{1}_\N{2048}_\N{10}_\N{1}_\V{base\_heapsort}}}
\L{\LB{\V{sim}\-\V{bpred}_\-\V{bpred}_\N{2$\ell$ev}_\-\V{bpred}:\N{2$\ell$ev}_\N{1}_\N{4096}_\N{10}_\N{1}_\V{base\_heapsort}}}