### 2005-05-26 paper/docs/software_predictor.tex

\afterpage{
\sixplot{base_heapsort-0.eps}{cache4_heapsort-0.eps}{cache4_heapsort-1.eps}{cache_heapsort-0.eps}{cache_heapsort-1.eps}{}{base and cache heapsorts}

\sixplot{base_quicksort-0.eps}{base_quicksort9-0.eps}{cache_quicksort-0.eps}{multi_quicksort-0.eps}{sequential_multi_quicksort-0.eps}{}{base, cache and multi quicksorts and base quicksort9}
\sixplot{base_quicksort1-0.eps}{base_quicksort-0.eps}{base_quicksort5-0.eps}{base_quicksort7-0.eps}{base_quicksort9-0.eps}{}{base quicksort without a median, with median of 3, and with pseudo median of 5, 7 and 9}

\sixplot{algorithm_n-0.eps}{algorithm_n-1.eps}{algorithm_s-0.eps}{algorithm_s-1.eps}{}{}{Knuth's Mergesorts}
\sixplot{base_mergesort-0.eps}{base_mergesort-1.eps}{base_mergesort-2.eps}{base_mergesort-3.eps}{base_mergesort-4.eps}{base_mergesort-5.eps}{base mergesort}
\sixplot{double_tiled_mergesort-0.eps}{double_tiled_mergesort-1.eps}{double_tiled_mergesort-2.eps}{double_tiled_mergesort-3.eps}{double_tiled_mergesort-4.eps}{double_tiled_mergesort-5.eps}{double tiled mergesort - part 1}
\sixplot{double_tiled_mergesort-6.eps}{double_tiled_mergesort-7.eps}{double_multi_mergesort-0.eps}{double_multi_mergesort-1.eps}{double_multi_mergesort-2.eps}{double_multi_mergesort-3.eps}{double tiled mergesort - part 2, and double multi mergesort - part 1}
\sixplot{double_multi_mergesort-4.eps}{double_multi_mergesort-5.eps}{double_multi_mergesort-6.eps}{double_multi_mergesort-7.eps}{double_multi_mergesort-8.eps}{double_multi_mergesort-9.eps}{double multi-mergesort - part 2}

\sixplot{gonnet_shellsort-0.eps}{gonnet_improved_shellsort-0.eps}{}{}{}{}{Shellsorts}
}