### 2005-05-27 paper/docs/report.tex
removed files not for paper removed extra bits

\documentclass[dvips]{report}
\usepackage{ifthen}
\usepackage{varioref}
\usepackage{listings}
\usepackage{color}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{float}
\usepackage{lgrind}

\usepackage[intlimits]{amsmath}
\usepackage{eufrak}
\usepackage[mathscr]{euscript}
\usepackage{afterpage}

\newenvironment{changemargin}{%
 \begin{list}{}{%
\setlength{\textwidth}{\paperwidth}
\setlength{\textheight}{\paperheight}
%\setlength{\oddsidemargin}{-1in}
%\setlength{\evensidemargin}{-1in}
\setlength{\topmargin}{-1in}
\setlength{\topsep}{-1in}%
	\setlength{\leftmargin}{-1.5in}%
	\setlength{\rightmargin}{-1.5in}%
	\setlength{\listparindent}{\parindent}%
	\setlength{\itemindent}{\parindent}%
	\setlength{\parsep}{\parskip}%
 }%
\centering%
\item[]%
}{\end{list}}

\newcommand{\cc}[1]{\texttt{#1}}
\newcommand{\n}[1]{\textit{#1}}
\newcommand{\code}[2]{\begin{figure}\lstinputlisting{code/#2}\caption{#1}\label{#1}\end{figure}}

\newcommand{\myscale}{1.2}
\newcommand{\plot}[7]{
% 3 figures first, then 3 page for them
\afterpage{
\setcounter{subfigure}{0}

\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-8em}
\begin{figure}[H]
\begin{changemargin}
\subfigure[Cycles per key - this was measured on a Pentium 4 using hardware performance counters. #2]
{\label{#1 cycles}\includegraphics[scale=\myscale]{plots/#1_-_cycles.eps}}
\subfigure[Instructions per key - this was simulated using SimpleScalar \cc{sim-cache}. #3]	
{\label{#1 instructions}\includegraphics[scale=\myscale]{plots/#1_-_Instructions.eps}}
\end{changemargin}
\vspace*{7em}
\caption{Simulated Instruction count and empiric cycle count for #1sort}
\label{Simulated Instruction count and empiric cycle count for #1sort}
\end{figure}
\newpage

\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-8em}
\begin{figure}[H]
\begin{changemargin}
\subfigure[Level 1 cache misses per key - this was simulated using SimpleScalar \cc{sim-cache}, simulating an 8KB data
cache with a 32 byte cache line and separate instruction cache. #4]
{\label{#1 level 1 misses}\includegraphics[scale=\myscale]{plots/#1_-_Level_1_misses.eps}}\\
\subfigure[Level 2 cache misses per key - this was simulated using SimpleScalar \cc{sim-cache}, simulating a 2MB data
and instruction cache and a 32 byte cache line. #5]
{\label{#1 level 2 misses}\includegraphics[scale=\myscale]{plots/#1_-_Level_2_misses.eps}}\\
\end{changemargin}
\vspace*{7em}
\caption{Cache Simulation results for #1sort}
\label{Cache Simulation results for #1sort}
\end{figure}
\newpage

\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-7em}
\begin{figure}[H]
\begin{changemargin}
\centering
\subfigure[Branches per key - this was simulated using \cc{sim-bpred}. #6]
{\label{#1 branches}\includegraphics[scale=\myscale]{plots/#1_-_Branches.eps}}
\subfigure[Branches misses per key - this was simulated using \cc{sim-bpred}, with bimodal and two-level adaptive
predictors. The simulated two-level adaptive predictor used a 10 bit branch history register which was XOR-ed with the
program counter. #7]
{\label{#1 branch misses}\includegraphics[scale=\myscale]{plots/#1_-_Branch_misses.eps}}
\end{changemargin}
\vspace*{6em}
\caption{Branch Simulation results for #1sort}
\label{Branch Simulation results for #1sort}
\end{figure}
}}

\newcommand{\smallscale}{.74}
\newcommand{\smallplot}[2]{
\setcounter{subfigure}{0}
\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-5em}
\begin{figure}[H]
\begin{changemargin}
\subfigure[Instructions per key] {\includegraphics[scale=0.74]{plots/#1_-_Instructions.eps}}\\
\subfigure[Level 1 cache misses per key]	{\includegraphics[scale=0.74]{plots/#1_-_Level_1_misses.eps}}
\subfigure[Level 2 cache misses per key]	{\includegraphics[scale=0.74]{plots/#1_-_Level_2_misses.eps}}
\subfigure[Branches per key]				{\includegraphics[scale=0.74]{plots/#1_-_Branches.eps}}
\subfigure[Branch misses per key]			{\includegraphics[scale=0.74]{plots/#1_-_Branch_misses.eps}}
\end{changemargin}
\vspace*{6em}
\caption{Simulation results for #2}
\label{Simulation results for #2}
\end{figure}
\newpage
}

\newcommand{\sixplot}[7]{
\afterpage{
\setcounter{subfigure}{0}
\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-5em}
\begin{figure}[H]
\begin{changemargin}
\subfigure[]{\includegraphics[scale=0.74]{../counters/#1}}
\ifthenelse{\equal{#2}{}}{}{\subfigure[]{\includegraphics[scale=0.74]{../counters/#2}}}
\ifthenelse{\equal{#3}{}}{}{\subfigure[]{\includegraphics[scale=0.74]{../counters/#3}}}
\ifthenelse{\equal{#4}{}}{}{\subfigure[]{\includegraphics[scale=0.74]{../counters/#4}}}
\ifthenelse{\equal{#5}{}}{}{\subfigure[]{\includegraphics[scale=0.74]{../counters/#5}}}
\ifthenelse{\equal{#6}{}}{}{\subfigure[]{\includegraphics[scale=0.74]{../counters/#6}}}
\end{changemargin}
\vspace*{6em}
\caption{Branch prediction performance for #7. All graphs use the same scale.}
\label{Branch prediction performance for #7}
\end{figure}
\newpage
}}

\lstset{language=c,tabsize=3,basicstyle=\small,showspaces=false,showtabs=false,showstringspaces=false,stepnumber=1,numberstyle=\small,frame=tblr}

\long\def\symbolfootnote[#1]#2{\begingroup\def\thefootnote{\fnsymbol{footnote}}\footnote[#1]{#2}\endgroup}

%\topmargin 0.2cm
%\textwidth 16cm
%\textheight 21cm
%\oddsidemargin 0.0cm
%\evensidemargin 0.0cm

\title{Fast Sorting on Modern Computers}
\author{Paul Biggar, David Gregg}
\date{}

\bibliographystyle{these}

\hyphenation{La-Marca}

%TODO make the page wider
%TODO aspell in UK english

% ----------------------------------------------------------------------------------
% END OF PREAMBLE
% ----------------------------------------------------------------------------------
\begin{document}

%\maketitle

\begin{abstract}
This project attempts to create and analyse sorts designed to take advantage of
features of modern processors. The work on cache-conscious sorting by LaMarca
and Ladner is the starting point; this is implemented and analysed using a
hardware simulator. Results relating to a variety of cache and branch predictor
configurations are obtained and analysed for a variety of sorts, beginning with
insertion and selection sorts, bubblesort and shakersort, and continuing to
heapsort, mergesort, quicksort and radixsort. Actual and expected results are
compared together and to the results of LaMarca's research.
\end{abstract}

\tableofcontents
\listoffigures

\setlength{\parskip}{1em}
\setlength{\parindent}{0ex}

\chapter{Introduction}

\input intro.tex

\chapter{Background Information}

\input background.tex

\chapter{Tools and Method}

\input method.tex

\chapter{Elementary Sorts}

\input ordernsquared.tex

\chapter{Heapsort}

\input heap.tex

\chapter{Mergesort}

\input merge.tex

\chapter{Quicksort}

\input quick.tex

\chapter{Radixsort}

\input radix.tex

\chapter{Shellsort}

\input shell.tex

\chapter{Conclusions}

\input conclusions.tex

\appendix

\chapter{Simulation Result Listing}
\input result_list.tex

%TODO \chapter{Software Predictor Result Listing}
%\input software_predictor.tex

% TODO the bugs are now fixed (actually, one of them was only mistakenly a bg :)
% the only remaining thing is the compulsory cache subtraction.
\chapter{Bug List}
\input bugs.tex

\bibliography{report} % param is the name of the .bib file
\nocite{*}

\end{document}