### 2005-06-10 fyp_report/slide7.tex

\begin{slide}
\begin{stitle}
Results
\end{stitle}

\begin{itemize}
\item Caches
	\begin{itemize}
	\item LaMarca's cache assertions are correct
	\plot{.82}{demo1}
	\pagebreak
	\plot{1.05}{demo2}
	\pagebreak
	\item Fully Associative can be worse than Direct Mapped
	\plot{.95}{demo3}
	\pagebreak
	\end{itemize}
\item Branch Predictors
	\begin{itemize}
	\item Two-level predictor needs large tables to rival bimodal
	\plot{.9}{demo4}
	\pagebreak
	\item Bimodal predictors perform the same regardless of size, for small functions, while two-level adaptive predictors don't
	\plot{.9}{demo4}
	\pagebreak
	\item Avoiding flow control misses is possible
		\begin{itemize}
		\item Radixsort
		\plotdouble{.51}{demo6-1}{demo6-2}
		\pagebreak
		\end{itemize}
	\item Some standard algorithms have very good prediction rates
		\begin{itemize}
		\item Insertion sort has N branch misses
		\plotdouble{.51}{demo7-1}{demo7-2}
		%insertion sort vs selection and bubble
		% put these two beside each other
		\pagebreak
		\item Binary search is worse than sequential search for small lists
		\plotdouble{.51}{demo8-1}{demo8-2}
		\pagebreak
		%binary vs sequential, next to each other again
		\item Radixsort has no comparison misses
		\plot{.95}{demo6-2}
		\pagebreak
		\end{itemize}
	\item Unsorted lists perform badly
		\begin{itemize}
		\item 4 / 8 heaps - Very complex to replace with sorted lists
		\plotdouble{.51}{demo9-2}{demo9-1}
		\pagebreak
		%compare 2 heap to 4/8 heap
		\end{itemize}
	\end{itemize}
\end{itemize}

\end{slide}