### 2005-06-14 docs/fyp_report/result_list.tex

\afterpage{
\smallplot{on2_insertsort_}{insertion sort}{}{}{}{}{}
\smallplot{on2_selectsort_}{selection sort}{}{}{}{}{}
\smallplot{on2_bubblesort_}{bubblesort}{}{}{}{}{}
\smallplot{on2_shakersort_}{shakersort}{}{}{}{}{}
%
\smallplot{base_heapsort_}{base heapsort}{}{}{}{}{}
\smallplot{cache4_heapsort_}{cache heapsort (4-heap)}{}{}{}{}{}
\smallplot{cache4_heapsort_small_cache_line}{cache heapsort (4-heap) (with 16-byte cache line)}{}{}{}{}{}
\smallplot{cache_heapsort_}{cache heapsort (8-heap)}{}{}{}{}{}
%
\smallplot{algorithm_n}{algorithm N}{}{}{}{}{}
\smallplot{algorithm_s}{algorithm S}{}{}{}{}{}
\smallplot{base_mergesort_}{base mergesort}{}{}{}{}{}
\smallplot{tiled_mergesort_}{tiled mergesort}{}{}{}{}{}
\smallplot{double_tiled_mergesort_}{double tiled mergesort}{}{}{}{}{}
\smallplot{multi_mergesort_}{multi-mergesort}{}{}{}{}{}
\smallplot{double_multi_mergesort_}{double multi-mergesort}{}{}{}{}{}
%
\smallplot{base_quicksort_}{base quicksort}{}{}{}{}{}
\smallplot{cache_quicksort_}{memory-tuned quicksort}{}{}{}{}{}
\smallplot{multi_quicksort_}{multi-quicksort (binary search)}{}{}{}{}{}
\smallplot{multi_quicksort_seq_}{multi-quicksort (sequential search)}{}{}{}{}{}
%
\smallplot{base_radixsort_}{radixsort}{}{}{}{}{}
}