### 2005-06-14 docs/fyp_report/simplescalar.tex

\label{SimpleScalar Install}

This section details how to install SimpleScalar on a Pentium 4 running RedHat
Linux 7. The reason for its inclusion is that the installation was difficult,
and solutions to problems encountered were not easily available on the
SimpleScalar website, nor were emails answered by the authors.

The primary guide this is based on is \cite{SimpleScalarInstall}. However, that
guide is more general than this.

\section{Download}
Go to \textit{http://www.simplescalar.com/tools}. The simulations in this
project used the standard packages provided on this page. Links to other
packages here were not used. Download simplesim-3v0d.tgz, simplesim-3v0d.tgz
and simplesim-3v0d.tgz. Unzip the tools. The following directories should be
present: binutils-2.5.2,  f2c-1994.09.27,  gcc-2.6.3,  glibc-1.09,
simplesim-3.0, ssbig-na-sstrix and sslittle-na-sstrix.

\section{Compile}
Run the following commands:
\begin{verbatim}
cd binutils-2.5.2
./configure --host=i386-redhat-linux --target=sslittle-na-sstrix 
	--with-gnu-as --with-gnu-ld --prefix=/home/ciaran/simple

make
make install
cd ../simplesim-3.0
make
cd ../gcc-2.6.3
./configure --host=i386-redhat-linux --target=sslittle-na-sstrix 
	--with-gnu-as --with-gnu-ld --prefix=/home/ciaran/simple
make LANGUAGES=c
\end{verbatim}

The following error may occur at this point:
\begin{verbatim}
cccp.c:194: conflicting types for `sys_errlist'
/usr/include/stdio.h:554: previous declaration of `sys_errlist'
\end{verbatim}

Comment out line 194 of \texttt{cccp.c}.

This error then occurs:
\begin{verbatim}
gcc -c -DCROSS_COMPILE -DIN_GCC -DPOSIX 
	-g -I. -I. -I./config sdbout.c
sdbout.c:57:18: syms.h: No such file or directory
\end{verbatim}

This can be resolved\footnote{Credit for this is due to
http://moss.csc.ncsu.edu/~mueller/rt/rt02/g1/sim2install.txt} by changing line
57 of \texttt{sdbout.c} from \verb!#include <syms.h>! to \verb!#include "gsyms.h"!.

\begin{verbatim}
bc-typecd.def: In function `bc_emit_instruction':
bc-typecd.def:17: `SFtype' is promoted to `double' 
					when passed through `...'
bc-typecd.def:17: (so you should pass `double' not 
					`SFtype' to `va_arg')
\end{verbatim}

In line 17 of \texttt{bc-typecd.def}, change \texttt{SFtype} to \texttt{DFtype}.

\begin{verbatim}
gcc.c:172: conflicting types for `sys_errlist'
/usr/include/stdio.h:554: previous declaration of `sys_errlist'
\end{verbatim}

Comment out line 172 of \texttt{gcc.c}

The tools should now be compiled. For speed, the architecture type could be
changed to \texttt{i686} and the \texttt{-O3} flag could be set.