### 2005-06-14 docs/fyp_report/slide2.tex

\begin{slide}

\begin{stitle}
Purpose
\end{stitle}

\begin{itemize}
\item Implement and test LaMarca's cache-conscious sorts
\begin{itemize}
\item expect same results as LaMarca's
\end{itemize}
\item Measure cache and branch prediction results
\begin{itemize}
\item Fully Associative vs. Direct Mapped Caches
\item Bimodal vs. 2 Level Adaptive Predictors
\end{itemize}
\item Do changes to algorithms:
\begin{itemize}
\item cause different branch effects?
\item increase/decrease branch misses?
\item affect different predictors differently?
\item \bfseries{Why?}
\end{itemize}
\end{itemize}

\end{slide}