### 2005-06-14 docs/tech_report/docs/bugs.tex

\label{bugs}

\section{Results}
\label{bug3}
To reduce the costs of creating the array of random numbers, this time was
measured and subtracted from the results. While this works as expected for
instruction count, the effects on cache misses is less clear. Subtracting level
2 cache misses incurred during filling removes compulsory misses from the
equation. For shorter lists, the compulsory misses are completely removed.
For longer lists, those greater than the size of the cache, compulsory misses
come back, since the keys at the start of the array are flushed from the cache
by the time they begin to be sorted. LaMarca's work centers around
reducing conflict misses, however, so this is not necessarily a problem, just
something to be aware of.