### 2005-06-14 docs/tech_report/docs/code/sample_output.tex

% Remember to use the lgrind style

\Head{}
\File{sample\_output}{2004}{4}{28}{4:03}{4376}
\L{\LB{\V{sim}\-\V{bpred}:_\V{SimpleScalar}/\V{PISA}_\V{Tool}_\V{Set}_\V{version}_\N{3.0}_\V{of}_\V{August},_\N{2003.}}}
\L{\LB{\V{Copyright}_(\V{c})_\N{1994}\-\N{2003}_\V{by}_\V{Todd}_\V{M}._\V{Austin},_\V{Ph}.\V{D}._\V{and}_\V{SimpleScalar},_\V{LLC}.}}
\L{\LB{\V{All}_\V{Rights}_\V{Reserved}._\V{This}_\V{version}_\V{of}_\V{SimpleScalar}_\V{is}_\V{licensed}_\K{for}_\V{academic}}}
\L{\LB{\V{non}\-\V{commercial}_\V{use}.}\Tab{21}{\V{No}_\V{portion}_\V{of}_\V{this}_\V{work}_\V{may}_\V{be}_\V{used}_\V{by}_\V{any}_\V{commercial}}}
\L{\LB{\V{entity},_\V{or}_\K{for}_\V{any}_\V{commercial}_\V{purpose},_\V{without}_\V{the}_\V{prior}_\V{written}_\V{permission}}}
\L{\LB{\V{of}_\V{SimpleScalar},_\V{LLC}_(\V{info}@\V{simplescalar}.\V{com}).}}
\L{\LB{}}
\L{\LB{\V{sim}:_\V{command}_\V{line}:_\V{sim}\-\V{bpred}_\V{base\_heapsort454574856}_}}
\L{\LB{}}
\L{\LB{\V{sim}:_\V{simulation}_\V{started}_@_\V{Wed}_\V{Apr}_\N{28}_\N{04}:\N{07}:\N{08}_\N{2004},_\V{options}_\V{follow}:}}
\L{\LB{}}
\L{\LB{\V{sim}\-\V{bpred}:_\V{This}_\V{simulator}_\V{implements}_\V{a}_\V{branch}_\V{predictor}_\V{analyzer}.}}
\L{\LB{}}
\L{\LB{\C{}\#_-config}\Tab{30}{\#_load_configuration_from_a_file\CE{}}}
\L{\LB{\C{}\#_-dumpconfig}\Tab{30}{\#_dump_configuration_to_a_file\CE{}}}
\L{\LB{\C{}\#_-h}\Tab{24}{false_\#_print_help_message}\Tab{54}{\CE{}}}
\L{\LB{\C{}\#_-v}\Tab{24}{false_\#_verbose_operation}\Tab{54}{\CE{}}}
\L{\LB{\C{}\#_-d}\Tab{24}{false_\#_enable_debug_message}\Tab{54}{\CE{}}}
\L{\LB{\C{}\#_-i}\Tab{24}{false_\#_start_in_Dlite_debugger\CE{}}}
\L{\LB{\-\V{seed}}\Tab{28}{\N{1}_\C{}\#_random_number_generator_seed_(0_for_timer_seed)\CE{}}}
\L{\LB{\C{}\#_-q}\Tab{24}{false_\#_initialize_and_terminate_immediately\CE{}}}
\L{\LB{\C{}\#_-chkpt}\Tab{23}{\<null\>_\#_restore_EIO_trace_execution_from_\<fname\>\CE{}}}
\L{\LB{\C{}\#_-redir:sim}\Tab{23}{\<null\>_\#_redirect_simulator_output_to_file_(non-interactive_only)\CE{}}}
\L{\LB{\C{}\#_-redir:prog}\Tab{23}{\<null\>_\#_redirect_simulated_program_output_to_file\CE{}}}
\L{\LB{\-\V{nice}}\Tab{28}{\N{0}_\C{}\#_simulator_scheduling_priority\CE{}}}
\L{\LB{\-\V{max}:\V{inst}}\Tab{28}{\N{0}_\C{}\#_maximum_number_of_inst{'}s_to_execute\CE{}}}
\L{\LB{\-\V{bpred}}\Tab{24}{\V{bimod}_\C{}\#_branch_predictor_type_\{nottaken\|taken\|bimod\|2lev\|comb\}\CE{}}}
\L{\LB{\-\V{bpred}:\V{bimod}}\Tab{15}{\N{2048}_\C{}\#_bimodal_predictor_config_(\<table_size\>)\CE{}}}
\L{\LB{\-\V{bpred}:\N{2$\ell$ev}}\Tab{15}{\N{1}_\N{1024}_\N{8}_\N{0}_\C{}\#_2-level_predictor_config_(\<l1size\>_\<l2size\>_\<hist\_size\>_\<xor\>)\CE{}}}
\L{\LB{\-\V{bpred}:\V{comb}}\Tab{15}{\N{1024}_\C{}\#_combining_predictor_config_(\<meta\_table\_size\>)\CE{}}}
\L{\LB{\-\V{bpred}:\V{ras}}\Tab{27}{\N{8}_\C{}\#_return_address_stack_size_(0_for_no_return_stack)\CE{}}}
\L{\LB{\-\V{bpred}:\V{btb}}\Tab{15}{\N{512}_\N{4}_\C{}\#_BTB_config_(\<num\_sets\>_\<associativity\>)\CE{}}}
\L{\LB{}}
\L{\LB{}\Tab{2}{\V{Branch}_\V{predictor}_\V{configuration}_\V{examples}_\K{for}_\N{2}\-\V{level}_\V{predictor}:}}
\L{\LB{}\Tab{4}{\V{Configurations}:}\Tab{22}{\V{N},_\V{M},_\V{W},_\V{X}}}
\L{\LB{}\Tab{6}{\V{N}}\Tab{10}{\C{}\#_entries_in_first_level_(\#_of_shift_register(s))\CE{}}}
\L{\LB{}\Tab{6}{\V{W}}\Tab{10}{\V{width}_\V{of}_\K{shift}_\V{register}(\V{s})}}
\L{\LB{}\Tab{6}{\V{M}}\Tab{10}{\C{}\#_entries_in_2nd_level_(\#_of_counters,_or_other_FSM)\CE{}}}
\L{\LB{}\Tab{6}{\V{X}}\Tab{10}{(\V{yes}\-\N{1}/\V{no}\-\N{0})_\V{xor}_\V{history}_\V{and}_\V{address}_\K{for}_\N{2nd}_\V{level}_\V{index}}}
\L{\LB{}\Tab{4}{\V{Sample}_\V{predictors}:}}
\L{\LB{}\Tab{6}{\V{GAg}}\Tab{14}{:_\N{1},_\V{W},_\N{2}\5\V{W},_\N{0}}}
\L{\LB{}\Tab{6}{\V{GAp}}\Tab{14}{:_\N{1},_\V{W},_\V{M}_(\V{M}_\>_\N{2}\5\V{W}),_\N{0}}}
\L{\LB{}\Tab{6}{\V{PAg}}\Tab{14}{:_\V{N},_\V{W},_\N{2}\5\V{W},_\N{0}}}
\L{\LB{}\Tab{6}{\V{PAp}}\Tab{14}{:_\V{N},_\V{W},_\V{M}_(\V{M}_==_\N{2}\5(\V{N}+\V{W})),_\N{0}}}
\L{\LB{}\Tab{6}{\V{gshare}}\Tab{14}{:_\N{1},_\V{W},_\N{2}\5\V{W},_\N{1}}}
\L{\LB{}\Tab{2}{\V{Predictor}_{`}\V{comb}\S{}{'}_combines_a_bimodal_and_a_2\-level_predictor.}}
\L{\LB{}}
\L{\LB{}}
\L{\LB{}}
\L{\LB{sim:_**_starting_functional_simulation_w/_predictors_**}}
\L{\LB{}}
\L{\LB{sim:_**_simulation_statistics_**}}
\L{\LB{sim\_num\_insn}\Tab{28}{1045897_\#_total_number_of_instructions_executed}}
\L{\LB{sim\_num\_refs}\Tab{29}{264304_\#_total_number_of_loads_and_stores_executed}}
\L{\LB{sim\_elapsed\_time}\Tab{34}{1_\#_total_simulation_time_in_seconds}}
\L{\LB{sim\_inst\_rate}\Tab{23}{1045897.0000_\#_simulation_speed_(in_insts/sec)}}
\L{\LB{sim\_num\_branches}\Tab{29}{167628_\#_total_number_of_branches_executed}}
\L{\LB{sim\_IPB}\Tab{29}{6.2394_\#_instruction_per_branch}}
\L{\LB{bpred\_bimod.lookups}\Tab{29}{167628_\#_total_number_of_bpred_lookups}}
\L{\LB{bpred\_bimod.updates}\Tab{29}{167628_\#_total_number_of_updates}}
\L{\LB{bpred\_bimod.addr\_hits}\Tab{29}{139898_\#_total_number_of_address\-predicted_hits}}
\L{\LB{bpred\_bimod.dir\_hits}\Tab{29}{140006_\#_total_number_of_direction\-predicted_hits}}
\L{\LB{bpred\_bimod.misses}\Tab{30}{27622_\#_total_number_of_misses}}
\L{\LB{bpred\_bimod.jr\_hits}\Tab{31}{8551_\#_total_number_of_address\-predicted_hits_for_JRs}}
\L{\LB{bpred\_bimod.jr\_seen}\Tab{31}{8552_\#_total_number_of_JR{'}\SE{}\V{s}_\V{seen}}}
\L{\LB{\V{bpred\_bimod}.\V{jr\_non\_ras\_hits}.\V{PP}_\N{0}_\C{}\#_total_number_of_address-predicted_hits_for_non-RAS_JRs\CE{}}}
\L{\LB{\V{bpred\_bimod}.\V{jr\_non\_ras\_seen}.\V{PP}_\N{0}_\C{}\#_total_number_of_non-RAS_JR{'}s_seen\CE{}}}
\L{\LB{\V{bpred\_bimod}.\V{bpred\_addr\_rate}}\Tab{31}{\N{0.8346}_\C{}\#_branch_address-prediction_rate\CE{}}}
\L{\LB{\V{bpred\_bimod}.\V{bpred\_dir\_rate}}\Tab{30}{\N{0.8352}_\C{}\#_branch_direction-prediction_rate\CE{}}}
\L{\LB{\V{bpred\_bimod}.\V{bpred\_jr\_rate}}\Tab{29}{\N{0.9999}_\C{}\#_JR_address-prediction_rate\CE{}}}
\L{\LB{\V{bpred\_bimod}.\V{bpred\_jr\_non\_ras\_rate}.\V{PP}_\<\V{error}:_\V{divide}_\V{by}_\N{0}\>_\C{}\#_non-RAS_JR_addr-pred_rate\CE{}}}
\L{\LB{\V{bpred\_bimod}.\V{retstack\_pushes}}\Tab{36}{\N{8554}_\C{}\#_total_number_of_address_pushed_onto_ret-addr_stack\CE{}}}
\L{\LB{\V{bpred\_bimod}.\V{retstack\_pops}}\Tab{34}{\N{8552}_\C{}\#_total_number_of_address_popped_off_of_ret-addr_stack\CE{}}}
\L{\LB{\V{bpred\_bimod}.\V{used\_ras}.\V{PP}}\Tab{32}{\N{8552}_\C{}\#_total_number_of_RAS_predictions_used\CE{}}}
\L{\LB{\V{bpred\_bimod}.\V{ras\_hits}.\V{PP}}\Tab{32}{\N{8551}_\C{}\#_total_number_of_RAS_hits\CE{}}}
\L{\LB{\V{bpred\_bimod}.\V{ras\_rate}.\V{PP}}\Tab{27}{\N{0.9999}_\C{}\#_RAS_prediction_rate_(i.e.,_RAS_hits/used_RAS)\CE{}}}
\L{\LB{}}