### 2005-06-15 docs/tech_report/result_list.tex

\label{result_list}
\afterpage{
\smallplot{on2_insertsort}{insertion sort}{}{}{}{}{}
\smallplot{on2_selectsort}{selection sort}{}{}{}{}{}
\smallplot{on2_bubblesort}{bubblesort}{}{}{}{}{}
\smallplot{on2_bubblesort2}{bubblesort2}{}{}{}{}{}
\smallplot{on2_shakersort}{shakersort}{}{}{}{}{}
\smallplot{on2_shakersort2}{shakersort2}{}{}{}{}{}
%
\smallplot{base_heapsort}{base heapsort}{}{}{}{}{}
\smallplot{cache4_heapsort}{cache heapsort (4-heap)}{}{}{}{}{}
\smallplot{cache4_heapsort_small_cache_line}{cache heapsort (4-heap) (with 16-byte cache line)}{}{}{}{}{}
\smallplot{cache_heapsort}{cache heapsort (8-heap)}{}{}{}{}{}
%
\smallplot{algorithm_n}{algorithm N}{}{}{}{}{}
\smallplot{algorithm_s}{algorithm S}{}{}{}{}{}
\smallplot{base_mergesort}{base mergesort}{}{}{}{}{}
\smallplot{double_tiled_mergesort}{double tiled mergesort}{}{}{}{}{}
\smallplot{double_multi_mergesort}{double multi-mergesort}{}{}{}{}{}
%
\smallplot{base_quicksort}{base quicksort}{}{}{}{}{}
\smallplot{cache_quicksort}{memory-tuned quicksort}{}{}{}{}{}
\smallplot{multi_quicksort}{multi-quicksort (binary search)}{}{}{}{}{}
\smallplot{multi_quicksort_seq}{multi-quicksort (sequential search)}{}{}{}{}{}
%
\smallplot{base_radixsort}{base radixsort}{}{}{}{}{}
\smallplot{cache_radixsort}{cache radixsort}{}{}{}{}{}
%
\smallplot{gonnet_shellsort}{gonnet shellsort}{}{}{}{}{}
\smallplot{improved_gonnet_shellsort}{improved gonnet shellsort}{}{}{}{}{}
}