### 2005-06-15 docs/tech_report/software_predictor.tex

\afterpage{
\sixplot{base_heapsort_0.eps}{cache4_heapsort_0.eps}{cache4_heapsort_1.eps}{cache_heapsort_0.eps}{cache_heapsort_1.eps}{}{base and cache heapsorts}

\sixplot{base_quicksort_0.eps}{base_quicksort9_0.eps}{cache_quicksort_0.eps}{multi_quicksort_0.eps}{sequential_multi_quicksort_0.eps}{}{base, cache and multi quicksorts and base quicksort9}
\sixplot{base_quicksort1_0.eps}{base_quicksort_0.eps}{base_quicksort5_0.eps}{base_quicksort7_0.eps}{base_quicksort9_0.eps}{}{base quicksort without a median, with median of 3, and with pseudo median of 5, 7 and 9}

\sixplot{algorithm_n_0.eps}{algorithm_n_1.eps}{algorithm_s_0.eps}{algorithm_s_1.eps}{}{}{Knuth's Mergesorts}
\sixplot{base_mergesort_0.eps}{base_mergesort_1.eps}{base_mergesort_2.eps}{base_mergesort_3.eps}{base_mergesort_4.eps}{base_mergesort_5.eps}{base mergesort}
\sixplot{double_tiled_mergesort_0.eps}{double_tiled_mergesort_1.eps}{double_tiled_mergesort_2.eps}{double_tiled_mergesort_3.eps}{double_tiled_mergesort_4.eps}{double_tiled_mergesort_5.eps}{double tiled mergesort - part 1}
\sixplot{double_tiled_mergesort_6.eps}{double_tiled_mergesort_7.eps}{double_multi_mergesort_0.eps}{double_multi_mergesort_1.eps}{double_multi_mergesort_2.eps}{double_multi_mergesort_3.eps}{double tiled mergesort - part 2, and double multi mergesort - part 1}
\sixplot{double_multi_mergesort_4.eps}{double_multi_mergesort_5.eps}{double_multi_mergesort_6.eps}{double_multi_mergesort_7.eps}{double_multi_mergesort_8.eps}{double_multi_mergesort_9.eps}{double multi-mergesort - part 2}

\sixplot{gonnet_shellsort_0.eps}{gonnet_improved_shellsort_0.eps}{}{}{}{}{Shellsorts}
}