### 2005-07-12 docs/tech_report/result_list.tex
updated gnuplot scripts, added scripts for newer sorts updated docs to inculde new sorts in the result list fixed bw

\label{result_list}
\afterpage{
\smallplot{on2_insertsort}{insertion sort}{}{}{}{}{}
\smallplot{on2_selectsort}{selection sort}{}{}{}{}{}
\smallplot{on2_bubblesort}{bubblesort}{}{}{}{}{}
\smallplot{on2_bubblesort2}{improved bubblesort}{}{}{}{}{}
\smallplot{on2_shakersort}{shakersort}{}{}{}{}{}
\smallplot{on2_shakersort2}{improved shakersort}{}{}{}{}{}
%
\smallplot{base_heapsort}{base heapsort}{}{}{}{}{}
\smallplot{cache4_heapsort}{cache heapsort (4-heap)}{}{}{}{}{}
\smallplot{cache_heapsort}{cache heapsort (8-heap)}{}{}{}{}{}
%
\smallplot{algorithm_n}{algorithm N}{}{}{}{}{}
\smallplot{algorithm_s}{algorithm S}{}{}{}{}{}
\smallplot{base_mergesort}{base mergesort}{}{}{}{}{}
\smallplot{tiled_mergesort}{tiled mergesort}{}{}{}{}{}
\smallplot{multi_mergesort}{multi-mergesort}{}{}{}{}{}
\smallplot{double_tiled_mergesort}{double tiled mergesort}{}{}{}{}{}
\smallplot{double_multi_mergesort}{double multi-mergesort}{}{}{}{}{}
%
\smallplot{base_quicksort1}{base quicksort (no median)}{}{}{}{}{}
\smallplot{base_quicksort}{base quicksort (median-of-3)}{}{}{}{}{}
\smallplot{base_quicksort5}{base quicksort (pseudo-median-of-5)}{}{}{}{}{}
\smallplot{base_quicksort7}{base quicksort (pseudo-median-of-7)}{}{}{}{}{}
\smallplot{base_quicksort9}{base quicksort (pseudo-median-of-9)}{}{}{}{}{}
\smallplot{cache_quicksort}{memory-tuned quicksort}{}{}{}{}{}
\smallplot{multi_quicksort}{multi-quicksort (binary search)}{}{}{}{}{}
\smallplot{multi_quicksort_seq}{multi-quicksort (sequential search)}{}{}{}{}{}
%
\smallplot{base_radixsort}{base radixsort}{}{}{}{}{}
\smallplot{cache_radixsort}{cache radixsort}{}{}{}{}{}
\smallplot{aligned_cache_radixsort}{aligned cache radixsort}{}{}{}{}{}
%
\smallplot{shellsort}{shellsort}{}{}{}{}{}
\smallplot{improved_shellsort}{improved shellsort}{}{}{}{}{}
}