### 2005-07-31 docs/tech_report/code/cycles_script.tex
done filename renaming inside files

% Remember to use the lgrind style

\Head{}
\File{cycles\_script}{2004}{4}{29}{0:06}{773}
\L{\LB{\K{set}_\V{key}_\V{top}_\V{left}_\V{Left}_\V{reverse}}}
\L{\LB{\K{set}_\V{xrange}_[\N{4096}:\N{4194304}]}}
\L{\LB{\K{set}_\V{yrange}_[\N{0}:]}}
\L{\LB{\K{set}_\V{xtics}_(\N{8192},\N{16384},\N{32768},\N{65536},\N{131072},\N{262144},\N{524288},\N{1048576},\N{2097152})}}
\L{\LB{\K{set}_\V{format}_\V{x}_\S{}\3\%8.0f\3\SE{}}}
\L{\LB{\K{set}_\V{logscale}_\V{axis}}}
\L{\LB{\K{set}_\V{data}_\V{style}_\V{linespoints}}}
\L{\LB{\K{set}_\V{term}_\V{postscript}_\V{eps}}}
\L{\LB{\K{set}_\V{output}_\S{}{'}.\,./docs/plots/all\_cycles.eps{'}\SE{}}}
\L{\LB{\K{set}_\V{xlabel}_\S{}{'}set_size_in_keys{'}\SE{}}}
\L{\LB{\K{set}_\V{ylabel}_\S{}{'}Cycles_per_key{'}\SE{}}}
\L{\LB{\V{plot}_\2}}
\L{\LB{\S{}{'}processed\_cycles\_data{'}\SE{}_\V{using}_\N{1}:\N{5}_\V{title}_\S{}{'}base_quicksort{'}\SE{},\2}}
\L{\LB{\S{}{'}processed\_cycles\_data{'}\SE{}_\V{using}_\N{1}:\N{6}_\V{title}_\S{}{'}memory_tuned_quicksort{'}\SE{},\2}}
\L{\LB{\S{}{'}processed\_cycles\_data{'}\SE{}_\V{using}_\N{1}:\N{7}_\V{title}_\S{}{'}multi_quicksort{'}\SE{},\2}}
\L{\LB{\S{}{'}processed\_cycles\_data{'}\SE{}_\V{using}_\N{1}:\N{10}_\V{title}_\S{}{'}base_mergesort{'}\SE{},\2}}
\L{\LB{\S{}{'}processed\_cycles\_data{'}\SE{}_\V{using}_\N{1}:\N{14}_\V{title}_\S{}{'}double_aligned_mergesort{'}\SE{},\2}}
\L{\LB{\S{}{'}processed\_cycles\_data{'}\SE{}_\V{using}_\N{1}:\N{16}_\V{title}_\S{}{'}double_multi\-mergesort{'}\SE{},\2}}
\L{\LB{\S{}{'}processed\_cycles\_data{'}\SE{}_\V{using}_\N{1}:\N{9}_\V{title}_\S{}{'}base_radixsort{'}\SE{}}}