### 2005-07-31 docs/tech_report/software_predictor.tex

\afterpage{
\sixplot{base_heapsort_0.eps}{memory_tuned4_heapsort_0.eps}{memory_tuned4_heapsort_1.eps}{memory_tuned_heapsort_0.eps}{memory_tuned_heapsort_1.eps}{}{base and memory-tuned heapsorts}

\sixplot{base_quicksort_0.eps}{base_quicksort9_0.eps}{memory_tuned_quicksort_0.eps}{multi_quicksort_0.eps}{sequential_multi_quicksort_0.eps}{}{base, memory-tuned and multi quicksorts and base quicksort (pseudo-median-of-9)}
\sixplot{base_quicksort1_0.eps}{base_quicksort_0.eps}{base_quicksort5_0.eps}{base_quicksort7_0.eps}{base_quicksort9_0.eps}{}{base quicksort without a median, with median of 3, and with pseudo-medians of 5, 7 and 9}
\sixplot{algorithm_n_0.eps}{algorithm_n_1.eps}{algorithm_s_0.eps}{algorithm_s_1.eps}{}{}{Knuth's mergesorts}
\sixplot{base_mergesort_0.eps}{base_mergesort_1.eps}{base_mergesort_2.eps}{base_mergesort_3.eps}{base_mergesort_4.eps}{base_mergesort_5.eps}{base mergesort}
\sixplot{double_aligned_mergesort_0.eps}{double_aligned_mergesort_1.eps}{double_aligned_mergesort_2.eps}{double_aligned_mergesort_3.eps}{double_aligned_mergesort_4.eps}{double_aligned_mergesort_5.eps}{double-aligned tiled mergesort - part 1}
\sixplot{double_aligned_mergesort_6.eps}{double_aligned_mergesort_7.eps}{double_aligned_multi_mergsort_0.eps}{double_aligned_multi_mergsort_1.eps}{double_aligned_multi_mergsort_2.eps}{double_aligned_multi_mergsort_3.eps}{double-aligned tiled mergesort - part 2, and double multi mergesort - part 1}
\sixplot{double_aligned_multi_mergsort_4.eps}{double_aligned_multi_mergsort_5.eps}{double_aligned_multi_mergsort_6.eps}{double_aligned_multi_mergsort_7.eps}{double_aligned_multi_mergsort_8.eps}{double_aligned_multi_mergsort_9.eps}{double-aligned multi-mergesort - part 2}

\sixplot{shellsort_0.eps}{improved_shellsort_0.eps}{}{}{}{}{Shellsorts}
}