### 2005-08-14 docs/tech_report/report.tex

\documentclass{report}
\usepackage{ifthen}
\usepackage{varioref}
\usepackage{listings}
\usepackage{color}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{float}
\usepackage{lgrind}

\usepackage[intlimits]{amsmath}
\usepackage{eufrak}
\usepackage[mathscr]{euscript}
\usepackage{afterpage}

% widens the margins for the graphs
\newenvironment{changemargin}{%
 \begin{list}{}{%
\setlength{\textwidth}{\paperwidth}
\setlength{\textheight}{\paperheight}
%\setlength{\oddsidemargin}{-1in}
%\setlength{\evensidemargin}{-1in}
\setlength{\topmargin}{-1in}
\setlength{\topsep}{-1in}%
	\setlength{\leftmargin}{-1.5in}%
	\setlength{\rightmargin}{-1.5in}%
	\setlength{\listparindent}{\parindent}%
	\setlength{\itemindent}{\parindent}%
	\setlength{\parsep}{\parskip}%
 }%
\centering%
\item[]%
}{\end{list}}

\newenvironment{spacedpars}{%for the abstract
\setlength{\parskip}{1em}
\setlength{\parindent}{0ex}
}

%\cc{x} is used for
% - unix commans or programs: gcc, sim-bpred, sim-cache, espresso, li, simple,
%   sim-fast, diff, valgrind, cachegrind, driver
% - code or keywords: less, Item, exch, if, while, unsigned int, int, qsort,
%   goto, do-while, THRESHOLD, partition, malloc, shellsort, improved shellsort
% - file names: main.c
% - #defines: UNIT_MAX
\newcommand{\cc}[1]{\texttt{#1}}

%\n{x} is used for
% - proper names: Atom, SimpleScalar, Papi, GnuPlot, ValGrind, glibc
% - technical names: unsigned integer, less-than, greater-than, in-place, out-of
%   place, Abstract Data Type, records, cache hierarchy, latency, hit, miss,
%   compulsory, cold-start, capacity, conflict, cache line, cache block, offset,
%   tag, virtaul addresses, bus address, temporal locality, locality, spatial
%   locality, taken, not taken, flow control branch, comparaitve branch, static,
%   semi-static, dynamic, 1-bit, bimodal, two-level adaptive, heap, children,
%   implicit heap, heapifying, sifting, fix-up, fix-down, sift-up, sift-down,
%   Repeated-adds, Remove-min, d-heap, fanout, natural, straight, unrolled,
%   Memory management unit, tiling, k-way merge, trashing divide and conquer,
%   multi-way partition, comparision based, counting based, increments
% - web addresses: http://www.gnuplot.com, http://valgrind.kde.org
% - sort names, when highlighted: algorithm S, algorithm N
% - branch predictor names: 'Creation, child promotion', etc
\newcommand{\n}[1]{\textit{#1}}

% to include a file as code
\newcommand{\code}[2]{\begin{figure}\lstinputlisting{code/#2}\caption{#1}\label{#1}\end{figure}}

% used for scaling graphs
\newcommand{\myscale}{1.2}

% plots graphs across 3 pages
\newcommand{\plot}[7]{
% 3 figures first, then 3 page for them
\afterpage{
\setcounter{subfigure}{0}

\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-8em}
\begin{figure}[H]
\begin{changemargin}
\subfigure[Cycles per key - this was measured on a Pentium 4 using hardware performance counters. #2]
{\label{#1 cycles}\includegraphics[scale=\myscale]{plots/#1_-_cycles.eps}}
\subfigure[Instructions per key - this was simulated using SimpleScalar \cc{sim-cache}. #3]	
{\label{#1 instructions}\includegraphics[scale=\myscale]{plots/#1_-_Instructions.eps}}
\end{changemargin}
\vspace*{7em}
\caption{Simulated instruction count and empiric cycle count for #1sort}
\label{Simulated instruction count and empiric cycle count for #1sort}
\end{figure}
\newpage

\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-8em}
\begin{figure}[H]
\begin{changemargin}
\subfigure[Level 1 cache misses per key - this was simulated using SimpleScalar
\cc{sim-cache}, simulating an 8KB data cache with a 32-byte cache line and
separate instruction cache. #4] {\label{#1 level 1
misses}\includegraphics[scale=\myscale]{plots/#1_-_Level_1_misses.eps}}\\
\subfigure[Level 2 cache misses per key - this was simulated using SimpleScalar
\cc{sim-cache}, simulating a 2MB data and instruction cache and a 32-byte cache
line. #5]
{\label{#1 level 2 misses}\includegraphics[scale=\myscale]{plots/#1_-_Level_2_misses.eps}}\\
\end{changemargin}
\vspace*{7em}
\caption{Cache simulation results for #1sort}
\label{Cache simulation results for #1sort}
\end{figure}
\newpage

\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-7em}
\begin{figure}[H]
\begin{changemargin}
\centering
\subfigure[Branches per key - this was simulated using \cc{sim-bpred}. #6]
{\label{#1 branches}\includegraphics[scale=\myscale]{plots/#1_-_Branches.eps}}
\subfigure[Branches misses per key - this was simulated using \cc{sim-bpred},
with bimodal and two-level adaptive predictors. The simulated two-level adaptive
predictor used a 10-bit branch history register which was XOR-ed with the
program counter. #7]
{\label{#1 branch misses}\includegraphics[scale=\myscale]{plots/#1_-_Branch_misses.eps}}
\end{changemargin}
\vspace*{6em}
\caption{Branch simulation results for #1sort}
\label{Branch simulation results for #1sort}
\end{figure}
}}

% the scaling for the smaller graphs
\newcommand{\smallscale}{.74}

% plots 6 images on 1 page
\newcommand{\smallplot}[2]{
\setcounter{subfigure}{0}
\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-5em}
\begin{figure}[H]
\begin{changemargin}
\subfigure[Cycles per key] {\includegraphics[scale=0.74]{plots/#1-_cycles.eps}}
\subfigure[Instructions per key] {\includegraphics[scale=0.74]{plots/#1_-_Instructions.eps}}
\subfigure[Level 1 cache misses per key]	{\includegraphics[scale=0.74]{plots/#1_-_Level_1_misses.eps}}
\subfigure[Level 2 cache misses per key]	{\label{#2 level 2 misses}\includegraphics[scale=0.74]{plots/#1_-_Level_2_misses.eps}}
\subfigure[Branches per key]				{\includegraphics[scale=0.74]{plots/#1_-_Branches.eps}}
\subfigure[Branch misses per key]			{\includegraphics[scale=0.74]{plots/#1_-_Branch_misses.eps}}
\end{changemargin}
\vspace*{6em}
\caption{Simulation results for #2}
\label{Simulation results for #2}
\end{figure}
\newpage
}

% plots 6 or less images on a page, using the counters
\newcommand{\sixplot}[7]{
\afterpage{
\setcounter{subfigure}{0}
\thispagestyle{empty}
\clearpage
\enlargethispage{14em}
\vspace*{-5em}
\begin{figure}[H]
\begin{changemargin}
\subfigure[]{\label{#7-1}\includegraphics[scale=0.74]{plots/counter_#1}}
\ifthenelse{\equal{#2}{}}{}{\subfigure[]{\label{#7-2}\includegraphics[scale=0.74]{plots/counter_#2}}}
\ifthenelse{\equal{#3}{}}{}{\subfigure[]{\label{#7-3}\includegraphics[scale=0.74]{plots/counter_#3}}}
\ifthenelse{\equal{#4}{}}{}{\subfigure[]{\label{#7-4}\includegraphics[scale=0.74]{plots/counter_#4}}}
\ifthenelse{\equal{#5}{}}{}{\subfigure[]{\label{#7-5}\includegraphics[scale=0.74]{plots/counter_#5}}}
\ifthenelse{\equal{#6}{}}{}{\subfigure[]{\label{#7-6}\includegraphics[scale=0.74]{plots/counter_#6}}}
\end{changemargin}
\vspace*{6em}
\caption[Branch prediction performance for #7]{Branch prediction performance for #7. All graphs use the same scale.}
\label{Branch prediction performance for #7}
\end{figure}
\newpage
}}

% settings for the code listings
\lstset{language=c,tabsize=3,basicstyle=\small,showspaces=false,showtabs=false,showstringspaces=false,stepnumber=1,numberstyle=\small,frame=tblr}

% i actually have no idea what this does, and no memory of writing it
\long\def\symbolfootnote[#1]#2{\begingroup\def\thefootnote{\fnsymbol{footnote}}\footnote[#1]{#2}\endgroup}

%\topmargin 0.2cm
%\textwidth 16cm
%\textheight 21cm
%\oddsidemargin 0.0cm
%\evensidemargin 0.0cm

\bibliographystyle{these}

% how to hyphenate words which arent standard
\hyphenation{La-Marca}

%TODO make the page wider
%TODO aspell in UK english
%TODO sometimes I say 3 sorts, sometimes 4. It's now 5.
%TODO do a final check for use of the word I
%TODO final check for the word 'project'

% ----------------------------------------------------------------------------------
% END OF PREAMBLE
% ----------------------------------------------------------------------------------
\begin{document}

\begin{titlepage}
       \begin{center}
				\vspace*{40mm}
                \Huge Sorting in the Presence of Branch Prediction and Caches\\
                \vspace{1em}
                \Large Fast Sorting on Modern Computers\\
                \vspace{2em}
                \Large Paul Biggar \hspace{1cm} David Gregg\\
				\vspace{2em}
				\large Technical Report TR-CS-2005-XX\\
                \large Department of Computer Science,\\
				\large University of Dublin, Trinity College,\\
				\large Dublin 2, Ireland.
        \end{center}
\end{titlepage}

\begin{spacedpars}
\begin{abstract}
Sorting is one of the most important and studied problems in computer
science. Many good algorithms exist which offer various trade-offs in
efficiency, simplicity and memory use. However most of these
algorithms were discovered decades ago at a time when computer
architectures were much simpler than today. Branch prediction and
cache memories are two developments in computer architecture that have
a particularly large impact on the performance of sorting algorithms.

This report describes a study of the behaviour of sorting algorithms
on branch predictors and caches. Our work on branch prediction is
almost entirely new, and finds a number of important results. In
particular we show that insertion sort causes the fewest branch
mispredictions of any comparison-based algorithm, that optimizations,
such as the choice of the pivot in quicksort, can have a large impact
on the predictability of branches, and that advanced two-level branch
predictors are usually worse at predicting branches in sorting
algorithms than simpler branch predictors. In many cases it is
possible to draw links between classical theoretical analyses of
algorithms and their branch prediction behaviour.

The other main work described in this report is an analysis of the
behaviour of sorting algorithms on modern caches. Over the last decade
there has been considerable interest in optimizing sorting algorithms
to reduce the number of cache misses. We experimentally study the
cache performance of both classical sorting algorithms, and a variety
of cache-optimized algorithms, primarily those proposed by LaMarca and
Ladner. Our experiments cover a much wider range of algorithms than
other work, including the $O(N^2)$ sorts and shellsort, all within a
single framework. We discover a number of new results, particularly
relating to the branch prediction behaviour of cache-optimized sorts.

We also developed a number of other improvements to the algorithms,
such as removing the need for a sentinel in heapsort. Overall, we
found that a cache-optimized radix sort was the fastest algorithm for
all large arrays; the absence of comparison branches means that the
algorithm causes almost no branch mispredictions.
\end{abstract}
\end{spacedpars}

\tableofcontents
\listoffigures

\setlength{\parskip}{1em}
\setlength{\parindent}{0ex}

\chapter{Introduction}

\input intro.tex

\chapter{Background Information}

\input background.tex

\chapter{Tools and Method}

\input method.tex

\chapter{Elementary Sorts}

\input elementary.tex

\chapter{Heapsort}

\input heap.tex

\chapter{Mergesort}

\input merge.tex

\chapter{Quicksort}

\input quick.tex

\chapter{Radixsort}

\input radix.tex

\chapter{Shellsort}

\input shell.tex

\chapter{Conclusions}

\input conclusions.tex

\appendix

\chapter{Simulation Result Listing}
\input result_list.tex

%TODO \chapter{Software Predictor Result Listing}
%\input software_predictor.tex

% the only remaining thing is the compulsory cache subtraction.
\chapter{Bug List}
\input bugs.tex

\bibliography{report} % param is the name of the .bib file
\nocite{*}

\end{document}