### 2007-01-18 docs/branch_paper/paper.tex
Moved nicks paper into the branch_paper directory

\documentclass[acmtocl]{acmtrans2m}

\usepackage{graphicx}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newdef{definition}[theorem]{Definition}
\newdef{remark}[theorem]{Remark}

\newcommand{\comment}[1]{}

\title{An Experimental Study of Sorting and Branch Prediction}

\author{
PAUL BIGGAR\footnote[1]{PUT IN IRCSET GRANT DETAILS}, NICHOLAS NASH\footnotemark[1] KEVIN WILLIAMS\footnotemark[1] and DAVID GREGG
\\University of Dublin, Trinity College
}

\begin{abstract}
Sorting is one of the most important and well studied problems in Computer Science. Many good algorithms
are known which offer various trade-offs in efficiency, simplicity, memory use, and other factors. However,
these algorithms do not take into account features of modern computer architectures that significantly
influence performance. Caches and branch predictors are two such features, and while
there has been a significant
amount of research into the cache performance of general purpose sorting algorithms, 
there has been little research on their branch prediction properties. In this paper we empirically examine
the behaviour of the branches in all the most common sorting algorithms. 
We also consider the interaction of cache optimization on the predictability of the branches in these algorithms.
We find insertion sort to have the fewest branch mispredictions of any
comparison-based sorting algorithm, that bubble and shaker sort operate in a fashion which makes their
branches highly unpredictable, that the unpredictability of shellsort's branches improves
its caching behaviour and that several cache optimizations have little effect on mergesort's branch
mispredictions. We find also that optimizations to quicksort -- for example the choice of pivot -- have a strong
influence on the predictability of its branches. We point out a simple way of removing branch instructions from 
a classic heapsort implementation, and show also that unrolling a loop in a cache
optimized heapsort implementation improves the predicitability of its branches. Finally, we note
that when sorting random data two-level adaptive branch predictors are usually no better than simpler bimodal predictors. This is despite 
the fact
that two-level adaptive predictors are almost always superior to bimodal predictors in general.
\end{abstract}

\category{E.5}{Data}{Files}[Sorting/Searching]
\category{C.1.1}{Computer Systems Organization}{Processor Architectures, Single Data Stream Architectures}[Pipeline architectures]
\terms{Algorithms, Experimentation, Measurement, Performance}
\keywords{Sorting, Branch prediction}

\begin{document}

\begin{bottomstuff}
Corresponding author's address: David Gregg, Department of Computer
Science, University of Dublin, Trinity College, Dublin 2, Ireland. {\tt
David.Gregg@cs.tcd.ie}.
\end{bottomstuff}

\maketitle

\section{Motivation}
Classical analyses of algorithms make simplifying assumptions about
the cost of different machine instructions. For example, the RAM model used for 
establishing asymptotic bounds and Knuth's MIX 
machine code both make drastically simplifying assumptions about the cost of
machine instructions \cite{KnuthVol1_97}. 
More
recently researchers have recognized that on modern computers the cost
of accessing memory can vary dramatically depending on whether the
data can be found in the first-level cache, or must be fetched from a
lower level of cache or even main memory.  This has spawned a great
deal of research on cache-efficient searching and sorting \cite{Nyberg+94,Agarwal96,LaMarca96b,LaMarca96a,LaMarca97,Xiao+00,Rahman+01,Wickremesinghe+02}.

Another type of instruction whose cost can vary dramatically is the
conditional branch. Modern pipelined processors depend on \emph{branch
prediction} for much of their performance.  If the direction of a
conditional branch is correctly predicted ahead of time, the cost of
the conditional branch may be as little as the cost of, say, an
integer add instruction. If, on the other hand, the branch is
mispredicted the processor must flush its pipeline, and restart
from the correct target of the branch. This cost is typically a large
multiple of the cost of executing a correctly-predicted branch. For
example, Intel Pentium 4 processors \cite{Intel248966-010,Intel249438-01} have pipelines of up 
to 31 stages,
meaning that a branch misprediction will cost around 30 cycles.  In
contrast, on the same processor the cost of a memory access that
misses the first-level cache and hits the second level cache is 23
cycles. Fortunately, the branches in most programs are very
predictable, so branch mispredictions are usually rare. Indeed, prediction
accuracies of greater than 90\% are typical \cite{Uht+97} with the
best predictors.

The cost of executing branches is particularly important for sorting
because the inner-loops of most sorting algorithms consist of
comparisons of items to be sorted. Thus, the predictability of
these comparison branches is critical to the performance of
sorting algorithms. In this paper we study the predictability
of branches in most of the major sorting algorithms. 
We focus on the behaviour of the branches whose outcome depends on a
comparison of keys presented to the sorting algorithm in its input. Throughout this paper we refer to
such branches as \textit{comparison branches}. Branches associated with controlling simpler aspects 
of the control flow of the algorithms are of much less interest because they are generally almost perfectly predictable.
We also examine cache conscious variations of many of the algorithms,
and show how optimizations for the cache influence the level of branch
mispredictions.

\section{Branch prediction}
\label{branch_prediction}

Branches are a type of instruction used for defining the flow control of programs. In high level languages they result from 
the use of statements like \texttt{if}, \texttt{while} and their variants. Branches can be either \textit{taken}, indicating that the
address they provide is the new value for the program counter, or they can be \textit{not-taken}, in which case sequential execution continues
as though the branch had not been present.

Branch instructions pose a difficulty for pipelined 
processors. In a pipelined processor, while the program counter points to a particular instruction, a potentially large
number of subsequent instructions are in a partially completed state.
Until it is known whether a branch is taken or not-taken, it is not possible to know what the next instructions to be executed are.
If the processor cannot tell what these next instructions are, it cannot fill its pipeline.
To keep the utilization of the pipeline at a reasonable level, modern processors attempt to anticipate the outcome of 
branch instructions by employing \textit{branch predictors}. A good discussion of pipelining and its associated issues can
be found in Hennessy and Patterson \citeyear{HennessyPatterson90}.

There are several kinds of branch predictors: static, semi-static and dynamic. We use only dynamic predictors in our experiments, 
because they are most commonly used in real processors. The statistics in the following paragraphs are taken from those derived
by Uht \textit{et al} \citeyear{Uht+97}, over the SPECint92 benchmarking suite.

The simplest type of dynamic predictor is referred to 
as a \textit{1-bit} predictor. It keeps a table recording, at each entry, whether a particular branch was taken or not-taken. 
Using the table, it predicts 
that a branch
will go the same way as it went on its previous execution. 1-bit predictors achieve 77\% to 79\% accuracy.

We refer to a 2-bit dynamic predictor as a \textit{bimodal} predictor. A bimodal predictor operates in the same manner as a 1-bit predictor but each table entry
can be thought of  as maintaining a counter from 0 to 3. The counter decrements on each
taken branch (with the exception of when the count is 0), and increments on each not-taken branch (with the exception of when the count is 3). With
counts equal to 0 or 1 the next branch is predicted as taken, with counts of 2 or 3 the branch is predicted not-taken. 
When the counter is 0, we say the predictor is in the \textit{strongly taken} state, since the branch must be not-taken twice before the not-taken
direction will be predicted again. When the counter is 1 we say the predictor is in the \textit{taken} state. Similarly counts of 2 and 3 are referred
to as \textit{not-taken} and \textit{strongly not-taken} respectively. Bimodal predictors achieve 78\% to 89\% accuracy.

Finally, some branch predictors attempt to exploit correlations in branch outcomes to improve accuracy. A \textit{two-level adaptive} predictor maintains
a \textit{branch history register}. This register records the outcomes of a number of previous branch instructions. For example, a register contents 
of \texttt{110010} indicates that the 6 previous 
branch outcomes were taken, taken, not-taken, not-taken, taken and not-taken.
For each branch, this register is used to index a table of bimodal predictors.
The program counter can also be included in the index, either concatenated or XORed with the indexing register. Two-level adpative
predictors are accurate about 93\% of the time.

In all the above, due to the size of the predictor table branches can collide. That is, two 
different branches can map to the same table location, often reducing accuracy.

\section{Experimental Setup}

We used truly random data provided by Haahr \citeyear{Random} in all our experiments. We used ten chunks of random data containing $2^{22}$ (4194304) keys.
Where a particular experiment used only part of the keys in a chunk, it used the left-most keys.
Our results are averaged over these chunks of data.

Processors do not generally provide access to the behaviour of their caches or branch predictors. However, a software processor simulator can be used to 
provide measurements 
of cache and branch prediction behaviour.
Our branch prediction and caching results were generated using the SimpleScalar PISA simulator version 3 \cite{SimpleScalarTutorialv4}. 
We used \texttt{sim-cache} and 
\texttt{sim-bpred} to generate results for caching and branch prediction characteristics respectively.
We used a variety of cache configurations; generally speaking we used an 8 KB level 1 data cache, 8 KB level 1 instruction cache,
and a shared 2 MB instruction and data level 2 cache, all with 32 byte cache lines. 
We also experimented with both direct mapped and fully associative caches. When presenting results
where the cache configuration is important, we will describe the precise configuration used. We present
results here only for direct mapped caches. The results for fully associative caches are similar \cite{BiggarGregg05}.

For the measurements taken using SimpleScalar, we used power of two sized sets of random keys ranging from $2^{12}$ to $2^{22}$ keys in size.
This is with the exception of the quadratic time sorting algorithms, for which the maximum set size used was $2^{16}$ keys. This was done
because of the very significant time it takes such algorithms to sort large inputs.

It takes time to fill arrays, and this can distort the relationship between the results for small set sizes, and larger set sizes, 
for which this time is amortised. As a result, we measured this time, and subtracted it from our results.
However, this practice also distorts the results. It removes compulsory cache misses from the results, so long as the data fits in the cache. 
When the data does not fit in the cache, the keys from the start of the array are faulted, with the effect that capacity misses reoccur, 
which are not discounted.

For branch prediction we simulated both bimodal and two-level adaptive predictors. 
We used power of two sized tables for the branch predictors. 
We used tables containing between $2^{11}$ and $2^{14}$ predictors.
When
presenting branch prediction results, the predictor configurations will be described precisely.
SimpleScalar's \texttt{sim-bpred} provides total results over
all the branches in a program. It is often desirable to examine particular branches individually however.
In order to do so, we added our own simulations of branch predictors to the programs. For our 
own simulations we average the results of each branch over ten runs of random data containing $2^{22}$ keys.

Finally we used \textit{PapiEx} \cite{PapiExManual} to access performance counters on a Pentium 4 1.6 Ghz. This allows cycle counts (i.e. running times)
to be
determined for the sorting algorithms on a real processor.
The results presented from these tests are averaged over 1024 runs and use the system's random number generator to provide data. The data
ranged in size from $2^{12}$ to $2^{22}$ keys in size.
The processor we used had an 8-way associative 256KB level 2 
cache with 64-byte cache lines. The separate data and instruction level 1 caches are 4-way associative, 8KB in size and have 32-byte cache lines.
The Pentium 4 uses an unspecified form of dynamic branch prediction. A reference manual \cite{Intel248966-010} mentions a branch history
register, so it is probable that the scheme is similar to two-level adaptive.

\section{Elementary sorts}

We begin by analysing the branch prediction behaviour of a number
of common elementary sorting algorithms, each operating in
$O(n^2)$ time. The main practical advantage of these
algorithms is their simplicity, which generally allows them to
operate faster than algorithms with $O(n \lg n)$ time complexity
for small inputs. Indeed, as we will see later, optimized versions
of $O(n \lg n)$ sorts use calls to elementary sorts for small inputs.
Knuth \citeyear{KnuthVol3_98} provides a good exposition of these elementary
sorting algorithms (as well as basic versions of the sorting algorithms
we shall examine in later sections).

\subsection{Selection Sort}
\label{selection_sort}

Selection sort is a very simple quadratic time sorting algorithm which begins by sequentially searching
for the smallest key in its input and moving it into position. The algorithm 
continues by repeatedly searching for and then moving the smallest key in the unsorted
section of its input into the appropriate position. The inner-loop of selection sort,
which is run a total of $n - 1$ times on an input \texttt{a[0..n - 1]} is shown below

\begin{verbatim}
  min = i;
  for(j = i + 1; j < n; j++)
    if(a[j] < a[min]) min = j;
  swap(a[i], a[min]);
\end{verbatim}

It is easy to see that selection sort performs approximately $n^2/2$ comparisons
and exactly $n - 1$ exchanges. 
On the $j^{th}$ iteration of the inner-loop the comparison branch is taken
if \texttt{a[j]} is the minimum of \texttt{a[i..j]}, which occurs with
probability $1 / (j - i + 1)$\footnote{Throughout this paper, for the purpose of analysis we make
the usual assumption that the input to be sorted consists of a random permutation of $\lbrace1, .., n\rbrace$.}. 
The expected number of times the branch is taken
on the first iteration is thus given by $H_n - 1$ where 
$H_n = \sum_{i = 1}^n 1/i$ is the $n^{th}$ harmonic number \cite{KnuthVol1_97}.
Since $H_n \approx \ln n$, and $\sum_{i = 1}^n H_i = O(n \ln n)$ the comparison branch
will be highly predictable. As expected, Figure \ref{elementary_sorts}(a) shows that number of mispredictions
per key grows in a similar fashion to $\ln n$.

If the array is initially in sorted order, then the value of \texttt{min}
will never change in the inner-loop. Since this implies the comparison branch
is never taken, it will be almost perfectly predictable. Similarly, 
if the array is sorted in reverse order, the value of \texttt{min} must
change at every element. That is, the comparison branch 
is taken at every element, and the branch is therefore almost perfectly
predictable.  From the point of view of branch prediction, frequent
changes (preferably with no simple pattern) are the main indicator of
poor performance, rather than the absolute number of times that the
branch resolves in each direction. The worst case for selection sort
would be an array which is partially sorted in reverse order; the value of
\texttt{min} would change frequently, but not predictably so.
Figure \ref{elementary_sorts}(a) shows the behaviour of selection
sort's branches.

\subsection{Insertion Sort}
\label{insertion_sort}

Insertion sort is another simple algorithm with quadratic
worst case time. It works by iterating the inner-loop below
a total of $n - 1$ times over its input

\begin{verbatim}
   item = a[i];
   while(item < a[i - 1]) 
   {
     a[i] = a[i - 1];
     i--;
   }
   a[i] = item;
\end{verbatim}
On the $i^{th}$ iteration \texttt{a[0..i - 1]} is sorted, allowing
\texttt{a[i]} to be placed in the correct position as shown in
the inner-loop above. On average, insertion sort performs approximately
$n^2/4$ comparisons and $n^2/4$ assignments in its inner-loop. Thus insertion sort
performs around half as many branches as selection sort.

A very pleasing property of insertion sort is that
it generally causes just 
a single branch misprediction per key, as Figure \ref{elementary_sorts}(a) shows. Typically
the inner \texttt{while} loop exit causes this misprediction, when
the correct location for the current item has been found.
In fact, a static
predictor would perform just as well as the two predictors shown, since it
would mispredict exactly once per key.

\begin{figure}
\centering
\begin{tabular}{cc}
\includegraphics[width=0.5\textwidth]{plots/selection_insertion_sort_branches.eps} & \includegraphics[width=0.5\textwidth]{plots/bubble_shaker_sort_branches.eps} \\
(a) & (b) \\
\end{tabular}
\caption{(a) Shows the branch mispredictions for selection and insertion sort using bimodal and two-level adaptive predictors both having 4096 table entries.
It is noteworthy the simple bimodal predictor has essentially identical performance to the two-level adaptive predictors.
(b) Shows the branch mispredictions for bubble and shaker sort using bimodal and two-level adaptive predictors, again both having 4096 table entries.
For bubble sort the bimodal and two-level adaptive predictors have close to identical performance. However, for shaker sort the simple bimodal
predictor results in fewer mispredictions than the two-level adaptive predictor. We also note that as a result of 
the partial sorting performed by bubble and shaker sort they incur far more
branch mispredictions per key than other $O(n^2)$ sorting algorithms.
\texttt{sim-bpred} was used to simulate the branch predictors in both (a) and (b). The two-level adaptive predictors used 10-bit history registers.}
\label{elementary_sorts}
\end{figure}

\subsection{Bubble sort}
\label{bubble_sort}

Bubble sort is another simple sorting algorithm, although notoriously
inefficient. It works by iterating the loop shown below at most $n - 1$
times on an input \texttt{a[0..n - 1]}.

\begin{verbatim}
for(j = 0; j < n - i - 1; j++)
{
    if(a[j + 1] < a[j])
       swap(a[j + 1], a[j]);
}
\end{verbatim} 
In this code $i$ is the outer-loop iteration counter, and begins at 0 and counts
towards $n - 1$. Bubble sort uses
information yielded by the comparisons which selection
sort simply discards. After a full iteration of the inner-loop shown larger
elements have moved to the right, closer to their final location.

Shaker sort is a variation of bubble sort in which there are two inner-loops
which are alternated. One is the inner-loop shown above. The other
inner-loop scans right-to-left moving small elements to the left in the
same manner as moving larger elements to the right has just been described.
Although shaker sort has the same worst-case time as bubble sort, it is generally
more efficient in practice.

Figure \ref{elementary_sorts}(b) shows the number of branch mispredictions per key for
bubble and shaker sort. As the data set gets large, bubble sort causes nearly 10000
branch mispredictions per key. Shaker sort executes a lot fewer
branches, and so has fewer mispredictions, but the rate of
misprediction is similar.

To understand the high misprediction rate of bubble sort we define $Q^k_j$
as the probability that $j^{th}$ inner-loop comparison branch is taken on the $k^{th}$ 
outer-loop iteration, $1 \leq k < n$. It turns out that

\begin{equation}
\label{bubble_equation}
Q^k_j = 
\left\{ 
\begin{array}{ll}
  \frac{j}{j + k} & \quad \mbox{if $j \leq n - k$}\\
  0 & \quad \mbox{otherwise}\\ 
\end{array} 
\right. 
\end{equation}

\noindent
This is a result of the fact that the $j^{th}$ comparison branch of the inner-loop
is taken only if \texttt{a[j]} is not larger than all of \texttt{a[0..j - 1]}.
For this
to be the case, the $(j + 1)^{th}$ branch in the previous outer-loop iteration must have been taken,
since otherwise the previous iteration of the outer-loop left the maximum of \texttt{a[0..j - 1]} in \texttt{a[j]}.
The probability that the $(j + 1)^{th}$ branch of the
previous outer-loop iteration is taken is $Q^{k - 1}_{j + 1}$. Given that the $(j + 1)^{th}$ branch of
the previous outer-loop iteration was taken, the probability that \texttt{a[j]} is
not larger than all of \texttt{a[0..j - 1]} is $1 - 1/(j + 1)$, thus,

\[
Q^k_j = Q^{k-1}_{j+1}\left(1 - \frac{1}{j + 1} \right)
\]

\noindent
The bases $Q^{1}_j = 1 - 1 / (j + 1)$ for $1 \leq j < n$ and $Q^1_n = 0$ give the solution shown in Equation \ref{bubble_equation}.
It is easy to see that in the first outer-loop iteration ($k = 1$) the comparison branch is very likely to be taken, whereas in the last
outer-loop iteration ($k = n - 1$), the comparison branch is very unlikely to be taken. Moving between these two extremes causes the intermediate
branches to be unpredictable. The remainder of this section examines this phenomenon in more detail.

In order to analyse the behaviour of bubble sort when a bimodal predictor is used,
it is useful to make use of the \textit{steady state predictability} function. This function
generally gives a good estimate of the probability that a particular branch will be correctly predicted,
given that we know the probability that the branch will be taken. For a bimodal predictor this function is
given by

\begin{equation}
CP(p) = \frac{3p^2 - 3p + 1}{2p^2 - 2p + 1}
\label{CP_equation}
\end{equation}

\noindent
Where $p$ is the probability that the branch in question will be taken. Although the derivation of this
function is simple, we relegate it to the appendix because we wish
to focus mainly on our experimental results. $CP$ can be used to examine the predictability of the branches in bubble sort as the outer-loop iterates.
Bubble sort executes $n - k$ comparison branches on its $k^{th}$ outer-loop iteration. Thus,
using Equation \ref{bubble_equation} the average probability that the comparison branch is taken on the $k^{th}$ outer-loop iteration
is

\begin{eqnarray*}
Q^k_{avg} &=& \frac{1}{n - k} \sum_{j = 1}^{n - k} \frac{j}{j + k} \\
          &=& 1 + \frac{k}{n - k}(H_k - H_n)\\
%  &=& \frac{n - k - k(H_n - H_k)}{n - k}\\
         &\approx& 1 + \frac{k}{n - k}\ln\frac{k}{n}
\end{eqnarray*}

\noindent
Where $H_n$ denotes the $n^{th}$ harmonic number. An approximation to the predictability
of the comparison branch of bubble sort on the $k^{th}$ iteration is then given by $CP(Q^k_{avg})$.
Figure \ref{bubble_predictability}(a) plots this, showing how the predictability of the comparison branch varies
with the number of outer-loop iterations. Figure \ref{bubble_predictability}(b)
shows the measured number of correctly predicted branches as the outer-loop iterates, these results
correspond approximately with what was obtained analytically in Figure \ref{bubble_predictability}(a).

\begin{figure}
\centering
\begin{tabular}{cc}
\includegraphics[width=0.5\textwidth]{plots/bubble_predictability.eps} & \includegraphics[width=0.5\textwidth]{plots/bubble_sweep_misses.eps} \\
(a) & (b) \\
\end{tabular}
\caption{(a) Shows $CP(Q^k_{avg})$ as the sweep number, $k$, increases (i.e. as the outer-loop iterates). This gives a good approximation to the 
predictability of the comparison branch of bubble sort. (b) Shows the number of correctly predicted branches in bubble sort, as well as the number of executed branches.
Initially the number of correctly predicted branches is close to the number executed, however, the number of correctly predicted branches rapidly 
decreases as the sweep number increases. Gradually, the branches become predictable again. These trends correspond to the 
variation in predictability shown in (a). The results of (b)
were obtained using our own software simulated bimodal predictor averaged over a large number of arrays with 8192 elements. Note also that in (b) 
the version of bubble sort used does not terminate early if no swaps are performed for an entire execution of the inner-loop.  }
\label{bubble_predictability}
\end{figure}

Initially the branches
are quite predictable, however their predictability reduces quite rapidly as the outer-loop iterates.
The incremental movement of large keys to the right in bubble sort causes this degredation in the
predictability of its branches. As the data becomes close to being
fully sorted the predictability of the branches improves again. The partial sorting bubble sort performs allows it to potentially converge to 
fully sorted data in fewer outer-loop iterations than selection sort. We can
detect that the data is sorted early if the branch shown in the inner-loop above
is never taken for a whole iteration. However, this partial sorting also greatly increases
the number of branch mispredictions bubble sort incurs, substantially slowing it over selection sort or insertion sort.

\subsection{Remarks}

Our results clearly demonstrate the importance of branch
prediction for the elementary sorting algorithms. 
Our shaker sort implementation has fewer cache misses than our selection sort implementation, as Figure \ref{shaker_plots}(a) shows. 
It also has a lower instruction count than selection sort. However it is slower due to branch mispredictions, as Figure \ref{shaker_plots}(b) shows.
Figure \ref{elementary_sorts} shows
the contrast between the branch prediction performance of selection sort and bubble or shaker sort. Where
selection sort causes on average fewer than 12 mispredictions per key for a set
size of 65536 keys, bubble and shaker sort cause many thousands.
Insertion sort behaves even better than selection sort, causing
almost a single misprediction per key. Since insertion sort
also has the fewest cache misses it has much to recommend it
as an elementary sorting algorithm. It is bubble sort's
partial sorting behaviour that wreaks havoc with branch prediction,
adding further weight to Knuth's \citeyear{KnuthVol3_98} point of view that it
has very little to recommend it, besides a catchy name and some interesting theoretical problems it
poses.

Finally our results show that over these elementary sorting algorithms two-level adaptive branch predictors are no
better than simpler bimodal predictors. 
This is noteworthy, because two-level
predictors are similar to, or sigificantly more accurate than bimodal
predictors for almost all other types of branches \cite{Uht+97}.
In fact, a bimodal predictor out-performs a two-level adaptive predictor for shaker sort, as
is shown in Figure \ref{elementary_sorts}(b).

\begin{figure}
\centering
\begin{tabular}{cc}
\includegraphics[width=0.5\textwidth]{plots/elementary_cache1_misses.eps} & \includegraphics[width=0.5\textwidth]{plots/elementary_cycles.eps}\\
(a) & (b) \\
\end{tabular}
\caption{(a) Shows the level 1 cache misses of the elementary sorting algorithms. 
These results were obtained by using \texttt{sim-cache} to simulate an 8KB direct mapped 
level 1 cache with 32 byte cache lines. Note that there are no level 2 cache misses since the simulated 2nd
level cache is 2MB in size.
(b) Shows the number of cycles per key for the elementary sorting algorithms. 
Despite bubble and shaker sort having better cache performance than selection sort, the vast number of branch mispredictions 
(see Figure \ref{elementary_sorts}(b)) they cause make them less efficient than selection sort in practice.
These results were measured using hardware performance counters on a Pentium 4.}
\label{shaker_plots}
\end{figure}

\section{Shellsort}

Shellsort \cite{Shell59} was the first in-place sorting algorithm
with time complexity better than $O(n^2)$.
Algorithms like selection and bubble sort use each comparison
to resolve at most one inversion (an inversion is 
a pair of keys in the input which appear in the wrong order) in the input, and
it is well known that this restricts their average case running
time to be in $O(n^2)$. Shellsort uses each comparison to
resolve more than one inversion, giving better performance.

Shellsort uses a sequence of decreasing integers $h_1, h_2, \ldots, h_k$
with $h_k = 1$, called increments. 
On the $i^{th}$ of its $k$ iterations, shellsort performs
an insertion sort but treats \texttt{a[j]} and \texttt{a[j + $h_i$]} as though
they were adjacent (rather than a normal insertion sort, which treats \texttt{a[j]} 
and \texttt{a[j + 1]} as adjacent).

The choice of increments influences shellsort's asymptotic behaviour.
We used Gonnet's increments \cite{Gonnet+91} for our implementation, because
we found them to be efficient empirically. Using
Gonnet's increments $h_i = \lfloor \frac{5}{11} h_{i - 1} \rfloor$, with $h_1 = \lfloor \frac{5}{11}n \rfloor$ and the
final increment being 1 as usual.

We use a straightforward implementation of shellsort, except we perform the final iteration
(i.e. the iteration in which we perform a standard insertion sort because the increment is 1) separately. 
For this final insertion sort
we extract $\min(a[0], \ldots, a[9])$ as a sentinel, removing the need
for a bounds check in its inner-loop. This also simplifies
the bounds checks on the outer-most loop of the other iterations.

\subsection{Branch Prediction Results}

\begin{figure}
\centering
\begin{tabular}{cc}
\includegraphics[width=0.5\textwidth]{plots/shellsort_branch_misses.eps} & \includegraphics[width=0.5\textwidth]{plots/shell_heap_cache_misses.eps} \\
(a) & (b) \\
\end{tabular}
\caption{(a) Shows the average branches misses per key for shellsort using a bimodal and two-level adaptive predictor, both having 4096 table entries.
On this occasion, the two-level adaptive predictor out-performs the simpler bimodal predictor. \texttt{sim-bpred}
was used to simulate these branch predictors.
(b) Shows that the average number of cache misses per key of our basic heapsort implementation
is higher than for our basic shellsort implementation. Shellsort incurs only a small number of cache misses
because on average a key moves just 0.9744 places per iteration. These results use \texttt{sim-cache} to simulate
an 8KB direct mapped cache with 32 byte cache lines.}
\label{shellsort_results_figs}
\end{figure}

The behaviour of the comparison branches in shellsort is
simple to analyse. There are $\lfloor \log_\frac{11}{5} n \rfloor$ passes, with
each pass performing an insertion sort (using the current
increment). As we saw in Section \ref{insertion_sort} insertion
sort causes about one branch misprediction per key, thus
we expect about $n \lfloor \log_\frac{11}{5}n \rfloor$ mispredictions in total, and
about $\lfloor \log_\frac{11}{5}n \rfloor$ mispredictions per key.

Figure \ref{shellsort_results_figs}(a)
shows the branch mispredictions per key on average for our shellsort
implementations, with bimodal and two-level adaptive predictors.
The scale on the x-axis is logarithmic, and the plots are nearly
straight lines, as we would expect if the branch mispredictions per key
are given by $\lfloor \log_\frac{11}{5} n \rfloor$. For example,
the final measurement on the x-axis is for a set size of 4194304 and
here we observe from the plot approximately 18 misses per key on average,
and $\lfloor \log_\frac{11}{5} 4194304 \rfloor = 19$.

The comparison branch of the insertion sort inner-loop of shellsort are very unpredictable, unlike the 
nearly perfectly predictable branches of a standard insertion sort.
We measured a predictability of about 50\% on average for this branch using a simulated bimodal
predictor.
A static predictor which predicts all branches as taken would be ideal for predicting the branches of the
insertion sorts performed by shellsort.
We found that on average, using Gonnet's increments, a key moves 
0.9744 places per iteration (note that a move is of $h$ positions, 
where $h$ is the current increment). Thus, although the insertion
sort branch will still cause just a single misprediction per key, 
its outcome is not nearly as predictable as a standard insertion
sort, where keys on average are moved a larger number of times
per inner-loop iteration.

One might expect shellsort's cache performance to be poor, since it repeatedly accesses
distantly separated keys in its input while the increments are large. Each such access
could potentially cause a cache miss. However, the fact that on average 
keys move just 0.9744 places per iteration
implies that large numbers of references in distant parts
of the input (which could lead to cache misses) are not generally necessary.
In fact, the number of cache misses for our shellsort implementations, although significant,
is smaller than for our basic heapsort implementation as Figure \ref{shellsort_results_figs}(b)
shows. For our input sizes, which were up to $2^{22}$ keys, our shellsort implementations were on
average about 50\% faster than the best of our heapsort implementations (see Section \ref{heapsort} for
further details on our heapsort implementations).

\section{Heapsort}
\label{heapsort}

\begin{figure}
\centering
\begin{tabular}{cc}
\includegraphics[width=0.5\textwidth]{plots/heapsort_branch_counts.eps} & \includegraphics[width=0.5\textwidth]{plots/heapsort_branch_misses.eps}\\
(a) & (b) \\
\includegraphics[width=0.5\textwidth]{plots/heapsort_cycle_counts.eps} & \includegraphics[width=0.5\textwidth]{plots/heapsort_8heap_branches.eps} \\
(c) & (d) \\
\end{tabular}
\caption{Branch prediction and performance characteristics of our heapsort implementations. (a) Shows
the average number of branches per key measured using \texttt{sim-bpred}. (b) Shows the average branch mispredictions per key for tables of 4096
predictors, again measured using \texttt{sim-bpred}. The two-level adaptive (which used a 10-bit history register) and bimodal predictors have close to identical performance, and to avoid
cluttering the diagram only one instance of a two-level adaptive predictor is shown, the remainder of the predictors are bimodal. (c) Shows the cycle counts of the algorithms measured using
Pentium 4 hardware performance counters. Note that the Pentium 4 has 32 byte cache lines in its first level cache, and 64 byte cache lines in its
second level cache.
Finally (d) shows the predictability of the unrolled branches associated with an 8-heap, measured with our own software simulated bimodal predictor.}
\label{heapsort_results}
\end{figure}

Another well known general purpose sorting algorithm is heapsort \cite{Williams64}. Heapsort's running time is $O(n \lg n)$.
Heapsort begins by constructing a \textit{heap}: a binary tree in which every level except possibly the deepest is entirely filled, with the deepest level filled from the
left. In addition, a heap must also satisfy the \textit{heap property}: A key is associated with each node that is always at least as large as the keys 
associated with all its children nodes. The sequence of keys resulting from a pre-order traversal of a heap's nodes can be used to represent it unambiguously in an array.

An unordered array can be transformed into a heap in $O(n)$ time, using the approach of Floyd \citeyear{Floyd64}.
Given a node whose children are heaps but who may itself
violate the heap property, an iteration of Floyd's approach swaps the node with the larger of its children, and then repeats the same procedure 
with the subtree rooted at that child, until the heap property is satisfied or a leaf is reached. We refer to this as ``sifting down'' a node. 
To build the entire heap, we first sift-down the deepest nodes having children, followed by their
parents, and so on until we reach the root.

Once the heap is built, we gradually destroy it in order to sort. We iterate the code below $n - 1$ times.

\begin{verbatim}
    tmp = a[0];
    a[0] = a[--n];
    a[n] = tmp;    
    sift_down(a, 0, n);
\end{verbatim}

\noindent
On the $i^{th}$ iteration we swap the largest key, 
\texttt{a[0]}, with \texttt{a[n - i]}. 
After each swap, the heap property is restored by sifting down the new root of the heap (an operation with worst case time $O(\lg n)$). 
After iteration $i$ the heap contains $n - i$ keys, and is found in \texttt{a[0..n - i - 1]}. Meanwhile 
\texttt{a[n - i..n - 1]} contains the $i$ largest keys of the input from smallest to largest.

\subsection{Heapsort Variations}

For our base heapsort implementation, we used the approach described in the previous section, but removed a bounds check from the sift down procedure
by introducing a sentinel at the cost of no extra instructions. The following optimization reduced the instruction count by about 10\%
on average compared to an implementation of sift-down which uses a bounds check. We have not seen it described elsewhere. In a heap, exactly one node has a single left child, all 
other nodes have either two children
or zero children. On every sift-down, we compare each node which has children with each of its children. As a result of the node with a single child, on 
every sift-down we
must check that the node being compared to its children has a right child as well as a left child. To avoid this check while building the heap, we can just 
insert
a maximal extra node at the end of the heap. When destroying the heap however, every second iteration requires the insertion of 
this extra node. To avoid this the code for destroying the heap shown in the previous section is changed to

\begin{verbatim}
    tmp = a[0];
    a[0] = a[--n];
    sift_down(a, 0, n);
    a[n] = tmp;
\end{verbatim}

\noindent
By not modifying \texttt{a[n - 1]} until after sifting down,
it acts as a sentinel for sift-down. This removes the need for the bounds check without introducing any extra instructions.

Aside from this optimization, we also applied caching optimizations of LaMarca and Ladner \citeyear{LaMarca96b} to give several cache-optimized versions of the
algorithm. Firstly, the heap can be padded so that siblings
always reside in the same cache line. Another cache optimization is realised by abandoning Floyd's heap building method and simply repeatedly growing the
heap one array element at a time, sifting each new key upwards as needed. This greatly improves the temporal locality of the algorithm when the data
does not fit entirely within the cache.

To increase the utilization of cache lines (and hence require fewer cache lines to be
fetched) we can use a $k$-ary heap instead of the binary heap that has been assumed until now. When using a $k$-ary heap, every node having children (except perhaps one)
has $k$ children. When sifting up or down a $k$-ary heap, the maximum (or minimum, depending on the heap property being used) child must be found, which
requires $k - 1$ comparisons. Of course, as $k$ grows the sort degenerates to a selection sort. Note also that for a $k$-ary heap, the optimization described
above, which avoids bounds checking while sifting down cannot be applied. However, it is still more efficient to add a sentinel on every second iteration of the heap 
destruction (since this introduces
$O(n)$ operations whereas the bounds check in the inner-loop introduces $O(n \lg n)$ operations). For modest values of $k$ a performance improvement
can be expected over a simple binary heap \cite{LaMarca96b}. We experimented with 4-heaps and 8-heaps, combined with the aforementioned cache optimizations.

\subsection{Branch Prediction Results}
\label{heapsort_branch_results_text}

Figure \ref{heapsort_results}(a) shows the average number of branches per key for our heapsort implementations. We note that the cache-optimized heapsort
implementations execute fewer branches per key than our base heapsort implementation. Although these algorithms execute fewer branches
per key on average, they cause a larger total number of branch mispredictions than our base heapsort implementation. Indeed, Figure
\ref{heapsort_results}(b) shows that the base heapsort implementation incurs the fewest branch mispredictions of any of the algorithms. 
This is despite the fact that it is executing more branches on average.
Figure \ref{heapsort_results}(b) also shows that the simpler
bimodal predictors are generally no worse at predicting branches than the two-level adaptive predictors. With the exception of shellsort, this is in 
keeping with the trend we observed for 
the other sorting algorithms examined so far.

Notwithstanding the base heapsort implementation having the lowest misprediction rate,
Figure \ref{heapsort_results}(c) shows that the most efficient implementation on our hardware configuration to be the 8-heap variation of heapsort,
followed by the 4-heap variation.
This suggests the improved cache performance of these algorithms outweighs their increased level of branch mispredictions shown in 
Figure \ref{heapsort_results}(b).

When using a $k$-ary heap, finding the maximum (or minimum) child at a particular node causes the $i^{th}$ comparison branch to be taken
with probability $1/(1 + i)$. The comparisons behave just like those in selection sort (see Section \ref{selection_sort}), and are therefore quite
predictable.  For our 8-heap we unrolled the
loop to find the minimum child into 7 separate branches. 
Figure \ref{heapsort_results}(d) shows the behaviour of these 7 branches, for an unrolled loop.
As expected, the later branches are gradually taken less often on average, with a corresponding increase in their predictability.
Unrolling this loop
not only removes the loop over-head, but is also experimentally observed to improve the average prediction accuracy by about 2\%. 
The average steady state predicitability in the unrolled case, is given by

\[
\frac{1}{7}\sum^7_{i=1}CP\left(\frac{1}{1 + i}\right) \approx 0.724
\]

\noindent
Where $CP$ is the steady state predictability function introduced in Section \ref{bubble_sort}.
Without unrolling, the comparison branch is taken on average with probability $\frac{1}{7}\sum^7_{i=1}\frac{1}{i + 1}$, and its predictability
can be estimated as

\[
CP\left(\frac{1}{7}\sum^7_{i=1}\frac{1}{1 + i} \right) \approx 0.706
\]

\noindent
As is observed experimentally, the unrolled loop results in about a 2\% greater prediction accuracy.

\section{Mergesort}
\label{mergesort}

Mergesort \cite{KnuthVol3_98} is a another $O(n \lg n)$ sorting algorithm, which works by 
repeatedly merging pairs of sorted lists into a single sorted list of twice the length. These merges can
be done very simply in linear time and space. Mergesort's
input can be thought of as $n$ sorted lists of length one. To simplify the description, we 
assume the number of keys, $n$, is a power of two. Mergesort makes $\lg n$
passes over the input where the $i^{th}$ pass merges sorted lists of length $2^{i - 1}$ into 
sorted lists of length $2^i$.

\subsection{Mergesort Variations}
\label{mergesort_variations}

For our base
mergesort implementation we apply several optimizations.
Firstly, our mergesort alternates the merge
operation between two arrays to avoid unnecessary copying.
Secondly, a bitonic sort as described by Sedgewick \citeyear{Sedgewick02},
removes instructions from the inner-loop. A bitonic sort
sorts every second list in reverse, merging then works
from the two ends of the array containing the elements to
be merged, terminating when the indices cross. Since the indices
cannot go outside the bounds of the array unless they cross first, 
this eliminates the need for a bounds check in the inner-loop.
An insertion sort is also initially used to create
sorted lists of length four, rather than allowing mergesort to
sort these lists.

We also analysed the branch prediction behaviour in several
cache conscious mergesort variations, using the work of LaMarca and Ladner \citeyear{LaMarca97} as a reference.
Mergesort is not in-place\footnote{Worst case linear time in-place merging is possible but is generally inefficient in practice, see for example Katajainen 
\textit{et al} \citeyear{Katajainen+96}.}
and uses an auxiliary array 
for
its merge operation. With a cache of size $C$, the largest input we can sort in the cache therefore has size $C/2$. To avoid conflict misses, the
arrays should be allocated so that the cache line index of the start of the source array is $C/2$ from the cache line index of the start of the destination array. 
This can
be done for both the level 1 and level 2 cache.

Since each key is only used once per pass
in mergesort, if the input is larger than the cache then we will have no temporal re-use of data at all. To address this we can \textit{tile}
mergesort: cache sized blocks of the input are mergesorted, and then merged as usual. Combined with alignment this variation is referred to as
tiled mergesort. Tiling mergesort in this way does not fully address the problem of its poor temporal re-use, since the merging of the cache sized blocks in tiled 
mergesort will exhibit no temporal re-use and again have very poor
cache performance. To address this, a multi-way merge of the cache sized blocks can be used. A multi-way merge combines a number of sorted lists into a 
single sorted
list while reading through each of them only once. A heap can be used to help implement the multi-way merge efficiently. 
This multi-mergesort variation has better cache performance. We refer to this combination of alignment and tiling with multi-way merging
simply as multi-mergesort.

\subsection{Branch Prediction Results}
\label{mergesort_results_text}

Figure \ref{mergesort_branch_results}(a) shows the average number of branches per key executed by
our mergesort implementations. Up to $2^{18}$ keys can be sorted in the simulated 2 MB level 2 cache
(each key is 4 bytes, and mergesort is not in-place). As a result, when the number of keys being 
sorted reaches $2^{19}$ there is a dramatic increase in the number of branches executed by multi-mergesort. This
is because a multi-way merge begins to be used. Figure \ref{mergesort_branch_results}(b) 
shows that there is a corresponding increase in the number of branch
mispredictions caused by multi-mergesort as a result of the multi-way merge.

Many of the additional comparison branches introduced by
the multi-way merge are associated with a $k$-ary heap, and are therefore quite predictable, as we saw in
Section \ref{heapsort_branch_results_text}. In addition, two-level adaptive predictors outperform bimodal predictors on multi-mergesort's branches.
Figure \ref{mergesort_branch_results}(b) shows this. Both predictors have 4096 entry tables. Varying the table size
of the bimodal predictor between 2048 and 8192 entries did not result in any reduction in the number of branch mispredictions.
On the other hand, increasing the table size of the two-level adaptive predictor gave incremental
improvements in branch predictability. This suggests that the two-level adaptive predictor is
able to exploit correlations between branch outcomes which occur during the multi-way merge.

Figure \ref{mergesort_branch_results}(c) shows the simulated level 2 cache misses per key for our mergesort
implementations. It is noteworthy that although multi-mergesort has the best cache performance, its cache
performance is only slightly better than tiled mergesort. However, as the set size increases the difference in cache
performance will slowly become more pronounced, as the trend in Figure \ref{mergesort_branch_results}(c) suggests. 
Figure \ref{mergesort_branch_results}(d) shows the cycles per key measured for our mergesort implementations
running on actual hardware. These results show that our less cache conscious tiled mergesort implementation
out-performs multi-mergesort. This is likely because the set sizes are not large enough to make the superior
cache performance of multi-mergesort the dominating factor in the comparative performances. Instead, for the 
data sizes we used, multi-mergesort's increased instruction count and branch mispredictions
cause its performance to be inferior to tiled mergesort.

\begin{figure}
\centering
\begin{tabular}{cc}
\includegraphics[width=0.5\textwidth]{plots/mergesort_branches_per_key.eps} & \includegraphics[width=0.5\textwidth]{plots/mergesort_branch_misses.eps} \\
(a) & (b) \\
\includegraphics[width=0.5\textwidth]{plots/mergesort_level_2_misses.eps} & \includegraphics[width=0.5\textwidth]{plots/mergesort_cycle_counts.eps} \\
(c) & (d) \\
\end{tabular}
\caption{(a) Shows the branches per key executed by our mergesort variations. The sudden increase in the number of executed branches for multi-mergesort
results from the use of the multi-way merge when the data no longer fits within the cache. (b) Shows the number of mispredictions per key, all predictors
had 4096 table entries. Bimodal predictors were used for base and tiled mergesort. These results were generated using \texttt{sim-bpred}, the
two-level adaptive predictor had a 10-bit history register.
(c) Shows the level 2 cache misses per key, base mergesort has by far the worst cache behaviour, while multi-mergesort
has slightly better cache performance than tiled mergesort. These results were generated by using \texttt{sim-cache} to simulate a
direct mapped 2MB cache with 32-byte cache lines. Finally (d) shows the cycles per key of our algorithms measured using Pentium 4 hardware performance
counters. Tiled mergesort performs the best out of the algorithms, despite its inferior cache performance compared to multi-mergesort.}
\label{mergesort_branch_results}
\end{figure}

\section{Quicksort}

Quicksort \cite{Hoare62} is an algorithm which selects a key from its input
called the pivot and then partitions the other keys into two sets,
one set containing keys at most equal to the pivot and another
containing keys at least equal to the pivot. The keys of these sets should
appear respectively to the left and right of pivot in sorted order.
Therefore they can be sorted independently, each using quicksort.

\subsection{Quicksort Variations}

For our basic quicksort implementation, we make use of a number
of improvements suggested by Sedgewick \citeyear{Sedgewick78}. Firstly, we
use an explicit stack instead of recursing, to save space. Secondly, after
partitioning, we always quicksort the set with the smaller number of elements first. 
This reduces the worst case stack space to $O(\lg n)$ from $O(n)$.
Thirdly, inputs of less than a certain \texttt{THRESHOLD} (typically \texttt{THRESHOLD} is 10) in size are not quicksorted
at all. Instead, insertion sort is run as a post-pass over the output of quicksort.
The smallest of the first \texttt{THRESHOLD} keys is used as a sentinel for insertion sort.
Finally, we use median-of-3 pivoting, greatly reducing the chances of very unbalanced
partitioning and also removing the need for bounds checks in the inner-loop, thus
greatly reducing the instruction count.

Improvements such as those just described are found in most efficient quicksort
implementations. We also investigated the branch prediction behaviour of cache
conscious quicksort implementations. LaMarca and Ladner \citeyear{LaMarca97} suggest that instead of running
insertion sort as a post-pass over the quicksorted input, each time quicksort
is presented with an input of size less than \texttt{THRESHOLD} it immediately
insertion sorts it. Although this increases the instruction count, it is more
than compensated for by the resultant drop in cache misses. We refer to this
as memory-tuned quicksort.

When quicksort's input fits within a particular level of the cache quicksort
has good cache properties at that level. To combat a dramatic increase
in cache misses when the input does not fit in the cache, LaMarca and Ladner also suggest multi-quicksort. 
Multi-quicksort chooses a set of pivots and performs a multi-way partition which divides the input into
cache sized containers. Each container is then quicksorted. We leave
the precise details of performing a multi-way partition to elsewhere \cite{LaMarca96a,BiggarGregg05}, it is sufficient
to note that we can decide which container a key should be placed in by using either a sequential
or binary search.

\begin{figure}
\centering
\begin{tabular}{cc}
\includegraphics[width=0.5\textwidth]{plots/counter_base_quicksort__median_of_3__0.eps} & \includegraphics[width=0.5\textwidth]{plots/counter_memory_tuned_quicksort_0.eps} \\
(a) Basic quicksort & (b) Memory-tuned quicksort \\ 
\includegraphics[width=0.5\textwidth]{plots/counter_multi_quicksort__binary_search__0.eps} & \includegraphics[width=0.5\textwidth]{plots/counter_multi_quicksort__sequential_search__0.eps} \\
(c) Multi-quicksort (binary search) & (d) Multi-quicksort (sequential search)
\end{tabular}
\caption{Overview of branch prediction behaviour in our quicksort implementations. Every figure shows the behaviour of the $i$ and $j$ branches when
using a median-of-3 pivot. As described in Section \ref{quicksort_branch_results_text}, these branches are about 60\% biased and 64\% predictable when using
the median-of-3. In (a) the \texttt{median} branch is the combined results of the branches which compute the median-of-3 (these branches are also executed for (b), (c) and (d)).
Comparing (a) with (b), (c) and (d), we see that the \texttt{insertion} branch associated with its insertion sort is slightly less predictable than in 
the other variations. This is due to it running as a post-pass. Finally, comparing (c) with (d) we see that the binary search branches of (c), 
\texttt{binary left} and \texttt{binary right}, are very unpredictable compared to the \texttt{sequential} branch of (d).
}
\label{branch_prediction_quicksort}
\end{figure}

\begin{figure}
\begin{verbatim}
      pv = a[0];
      i = l, j = r + 1;
      while(true)
      {
              while(a[++i] < pv) ; // i-loop
              while(a[--j] > pv) ; // j-loop
              if(i >= j) break;
              swap(a[i], a[j]);
      }
      swap(a[0], a[j]);
\end{verbatim}
\caption{Quicksort's partition inner-loop. We refer to the inner while loops
as the $i$ and $j$ loops. We refer to their associated branches as
the $i$ and $j$ branches respectively.}
\label{partition_code}
\end{figure}

\subsection{Branch Prediction Results}
\label{quicksort_branch_results_text}

Figure \ref{branch_prediction_quicksort} shows the branch prediction results of our quicksort implementations.
In all figures, the most important branches are the $i$ and $j$ branches (see Figure \ref{partition_code}), since these 
are in quicksort's inner-loop, and are executed more often than any of the other branches (with the exception of the \texttt{sequential} branch
of Figure \ref{branch_prediction_quicksort}(d)).

The branch prediction behaviour of basic quicksort and memory-tuned quicksort are similar.
As can be seen in Figure \ref{branch_prediction_quicksort}(b) 
quicksort has a slightly lower branch misprediction rate on the \texttt{insertion} branch. This
is because a basic quicksort implementation, which performs insertion sort as a post-pass,
will generate branch mispredictions on keys chosen as pivots since they never require movement.
In practice, the over-all reduction in branch misprediction rate from 35\% in basic quicksort
to 34\% in memory-tuned quicksort is probably offset by the branch misprediction rate of the
outer-loop exit in the many insertion sorts executed by memory-tuned quicksort. However,
memory-tuned quicksort has better caching properties, and so from the point of view of both
branch prediction and caching behaviour this optimization is to be recommended.

Multi-quicksort results in a 19\% reduction in the number of iterations of the $i$ and $j$
loops of quicksort's partition inner-loop, 
though their misprediction rate remains the same at 37\%. Figure \ref{branch_prediction_quicksort}(c) shows branch behaviour
for multi-quicksort when using binary search. The binary search branches are highly unpredictable,
being incorrectly predicted 43\% of the time. If multi-quicksort uses sequential search, the
number of executed branches dramatically increases, but these
branches are correctly predicted 92\% of the time, as shown in Figure \ref{branch_prediction_quicksort}(d).

The implementations of multi-quicksort have the best cache performance of the quicksort variations, as Figure \ref{quicksort_plots}(a) shows.
However, the multi-quicksort implementations also introduce many extra instructions per key as Figure \ref{quicksort_plots}(b) shows. Naturally, the 
sequentially searched version introduces more extra instructions than the binary searched version. However, as described in the previous paragraph,
the branches introduced by binary search are highly unpredictable. As a result, multi-quicksort using binary search has more branch mispredictions than the
sequential searched version (see Figure \ref{quicksort_plots}(c)), consequently, despite having identical cache performance, binary searched multi-quicksort
performs worse than the sequentially searched multi-quicksort, as shown in Figure \ref{quicksort_plots}(d). Of course, on very large data sets (larger than the $2^{22}$ keys we experimented with), the 
reduced instruction count of binary searched multi-quicksort would eventually allow it to out-perform sequentially searched multi-quicksort.

As Figure \ref{quicksort_plots}(d) shows, our base quicksort implementation performs the best of all. It seems the improved cache performance of
multi-quicksort is not enough for it to out-perform base quicksort on the data set sizes we experimented with. Examining the trends of 
Figure \ref{quicksort_plots}(a), it seems likely that multi-quicksort would eventually out-perform base quicksort due to its improved cache performance.
The fact that base quicksort out-performs memory-tuned quicksort is very difficult to explain. In the simulations memory-tuned quicksort has the same
cache performance and instruction count, as shown in Figures \ref{quicksort_plots}(a) and \ref{quicksort_plots}(b). Also in the simulations, 
memory-tuned quicksort has slightly fewer branch mispredictions than base quicksort, as Figure \ref{quicksort_plots}(c) shows. Despite this, base 
quicksort slightly out-performs memory-tuned quicksort on our non-simulated hardware test as Figure \ref{quicksort_plots}(d) shows.

\begin{figure}
\centering
\begin{tabular}{cc}
\includegraphics[width=0.5\textwidth]{plots/quicksort_cache_misses.eps} & \includegraphics[width=0.5\textwidth]{plots/quicksort_instruction_counts.eps} \\
(a) & (b) \\ 
\includegraphics[width=0.5\textwidth]{plots/quicksort_branch_misses.eps} & \includegraphics[width=0.5\textwidth]{plots/quicksort_cycles.eps} \\
(c) & (d) \\
\end{tabular}
\caption{(a) Shows the 2nd level cache misses for our quicksort implementations. As is expected, the multi-quicksort implementations have the best
cache performance. These results were generated using \texttt{sim-cache} to simulate a 2MB direct-mapped cache with 32-byte cache lines.
(b) Shows the instruction counts of our quicksort implementations, when the data sets no longer fit within the cache the multi-quicksort implementations
show a large increase in instruction count, these results were also generated using SimpleScalar. (c) Shows the simulated branch mispredictions per key
for our quicksort implementations. All predictors use 4096 table entries and are bimodal predictors except where indicated. Multi-quicksort using
binary search is the only implementation which gives varying results from the use of a two-level adaptive predictor. These results were generated
using \texttt{sim-bpred}. (d) Shows the cycles per key for our quicksort implementations measured using Pentium 4 hardware performance counters.
}
\label{quicksort_plots}
\end{figure}

\subsection{Pivot Choice}
\label{quicksort_pivot_choice}

The choice of pivot element has a strong influence on the behaviour of the $i$ and 
$j$ branches (see Figure \ref{partition_code}). Clearly, if the pivot is chosen as the median of the input then
either branch is equally likely to be taken (i.e. the loop executes) as not taken (i.e.
the loop exits). In this case, the branches are also on average approximately 50\% predictable. 
However, on average over random input, with a fixed choice of pivot the $i$ and
$j$ branches are about 66\% biased (that is, they are taken about 66\% of the time). It is
straightforward to show why this is so. Assume that the data to be partitioned is a random permutation of $\lbrace 1, \ldots, n \rbrace$.
If the chosen pivot has rank $q$ then the $i$ branch will be taken with probability $(q - 1)/n$.
Moreover, the branch will be executed (but not necessarily taken) a total of $q$ times, since there are exactly $q - 1$ elements
to the left of the pivot. The average bias is given by averaging over all 
possible choices of pivot $q$, thus

\begin{equation}
B_{avg}^n = \frac{2}{n(n + 1)}\sum_{q = 1}^n \left(\frac{q - 1}{n} \right) q = \frac{2}{3} - \frac{2}{3n} 
\label{bias_equation}\end{equation}

\noindent
So the bias rapidly approaches 66\% as $n$ increases. With this bias, the $i$ and $j$ branches are about 71\% predictable. 
The fact that predictability of these branches exceeds their biases
may at first seem paradoxical. The reason the branches are more predictable than biased is because the bias 
is an average. The average predictability of the $i$ and $j$ branches can be obtained very simply from
Equation \ref{bias_equation} using the steady state predictability function, $CP$, introduced in Section \ref{bubble_sort}, thus

\[
P_{avg}^n = \frac{2}{n(n + 1)}\sum_{q = 1}^n CP\left( \frac{q - 1}{n} \right) q
\]

\noindent
As $n \rightarrow \infty$ we have $P_{avg} = \int_0^1 CP(q)\,dq = 3/2 - \pi/4$.
Thus we see $P_{avg}$ is about 0.71. This estimated average predictability is close to what is observed experimentally when using a fixed pivot
over random data, as the predictability entry of the first row of Table \ref{median_table} shows.

If the pivot is chosen as the median-of-3, then the $i$ and $j$ branches
are about 60\% biased in the taken direction, and about 64\% predictable. It is especially
noteworthy that although median-of-3 gives around a 14\% reduction in the total number of executed
comparison branches, there is actually a 6\% \textit{increase} in the total
number of branch mispredictions. As the pivot more closely approximates the median of the input,
there is a continuing trend of a reduction in the total number of executed branches, as well as reducing bias and predictability 
with an increasing total number of mispredictions, as Table \ref{median_table} shows.

\begin{table}
\begin{tabular}{|c|c|c|c|c|}
\hline
Median-of & Reduction In $\#$Branches & Bias    & Predictability & Misprediction Increase \\
\hline                                                                                   
1         & 0\%                       & 66.1\%  & 70.6\%         & 0\%                    \\
\hline
3         & 14\%                      & 60.0\%  & 64.3\%         & 6.1\%                  \\
\hline
5         & 16.5\%                    & 57.8\%  & 61.6\%         & 9.2\%                  \\                     
\hline
7         & 17.8\%                    & 56.3\%  & 59.9\%         & 11.8\%                 \\
\hline
9         & 19.6\%                    & 55.1\% & 58.3\%          & 14.3\%                 \\
\hline
\end{tabular}
\caption{The effect of choosing a pivot closer to the median on the prediction
of the $i$ and $j$ branches (the results are the same for both). As a higher order median is used the bias of the $i$ and $j$
branches is reduced, with a corresponding reduction in their predictability.
The reductions in the number of executed branches are relative to a fixed choice of pivot (i.e. median-of-1). 
The misprediction increases correspond to the total number
of branches mispredicted with a higher order median compared to a fixed choice of pivot.
These measurements were obtained 
from our own software simulations of a bimodal branch predictor. Note that the small number of branches
required to compute the median approximations are included in these results.}
\label{median_table} 
\end{table}

The use of the median-of-3 or higher order median approximation greatly reduces the
chances of the worst case behaviour of quicksort, and on average therefore results
in fewer comparison branches being executed. From an information theoretic
perspective, each comparison branch resolves more uncertainty, or yields
more information, and is hence harder to predict. This observation applies also
to the predictability of the branches used in the binary searched version of 
multi-quicksort versus the sequentially searched version. There is a trade-off
to be had between the penalty induced by unpredictable comparisons, versus the 
increased instruction count caused by making comparisons more predictable.

In work concurrent and independent of our own, Kaligosi and Sanders \citeyear{Kaligosi+06} 
have investigated the effect of the choice of pivot on the performance of quicksort. They found that
an artificially skewed pivot gives better performance than using a random pivot or the median-of-3,
due to the reduction in branch mispredictions. This performance increase is despite the increased instruction count
such an artificially skewed pivot causes, we describe this work in more detail in Section \ref{RelatedWork}.

\section{Radix Sort}
\label{Radixsort}

The purpose of this paper is to show the relevance of branch mispredictions to sorting. We now show how to develop a  
radix sort \cite{Friend56} implementation for 32-bit integers that, in all our tests, is more efficient than all other sorting algorithms presented in this paper. 
This is in part due to the tiny number of branch mispredictions incurred by radix sorting. It is also in part because radix sort 
operates in linear worst case time. However, in spite of its better asymptotic performance a ``step'' for our radix sort is quite costly, requiring
a key to be counted four times and copied four times, with the counting involving several layers of indirection. In contrast, a step for quicksort
is as simple as a comparison and a possible exchange every few steps.

The cache performance of radix sort is generally poor compared
to algorithms like mergesort or quicksort. In radix sort, each key is used just once each time it is in the cache, when the data to be sorted
is larger than the cache every key access will result in a cache miss. On the other hand, mergesort and quicksort have good temporal locality and use keys 
repeatedly before they are ejected from the cache.

\subsection{Memory-tuned Radix Sort}

Our basic radix sort implementation is a typical least significant digit radix sort. We operate in radix 256, and begin by counting the number of keys
with a particular least significant byte by indexing a table of 256 entries with the least significant bytes of the keys. The entries of this table are then summed to 
store the cumulative count of keys with a smaller least significant byte. Now each key is placed in an auxiliary array according to the value retrieved by 
using its least significant byte to index into the table of cumulative counts. Before the next iteration, the auxiliary array is written back to the 
original array and the sort can proceed on the next most significant byte. We use this implementation to compare against the simple caching improvements
we now suggest.

We can firstly remove the copying back and forth between the original and auxiliary array, by just alternating their roles on each step. Since we are
sorting 32 bit integers, the number of iterations is even and the data will end up sorted in its original location. If the number of steps was odd, an
additional copy would be required at the end. Since radix sort is out-of-place, we might consider aligning its two arrays as we did with mergesort (see
Section \ref{mergesort_variations}). However, radix sort does not access its arrays in the same regular fashion that mergesort does and so is unlikely to
benefit from such a change. Indeed, we implemented a double-aligned radix sort and did not observe any performance gains.
Finally, we use an improvement suggested by LaMarca and Ladner \citeyear{LaMarca97}, where the counting 
of the next most significant bytes is begun during their preceding iteration. We refer to this optimized version of radix sort as memory-tuned
radix sort.

In our basic radix sort implementation, the first array is accessed 3 times and the second array 4 times, in each of the four
iterations over the four bytes in the 32-bit integers. This gives 20 accesses in total.
In memory-tuned radix sort each key of the two arrays is accessed only nine times in total.
At the start, it traverses the original array once to count the least significant bytes. It then iterates four times, positioning
the keys based on the previously established counts. These four iterations cause
a total of eight accesses, because each array is accessed once per iteration.

\subsection{Results}

\begin{figure}
\centering
\begin{tabular}{cc}
\includegraphics[width=0.5\textwidth]{plots/radix_cycles.eps} & \includegraphics[width=0.5\textwidth]{plots/radix_versus_quick_level_2_misses.eps} \\
(a) & (b) \\
\includegraphics[width=0.5\textwidth]{plots/radix_branch_misses.eps} & \includegraphics[width=0.5\textwidth]{plots/radix_branches.eps} \\
(c) & (d) \\
\end{tabular}
\caption{(a) Shows the cycles per key of our radix sort implementations alongside our base quicksort implementation, measured using Pentium 4
hardware performance counters. (b) Shows the level 2 cache 
misses incurred by our radix sort implementations and a multi-quicksort implementation for a 2MB direct mapped cache with 32-byte cache lines. These
results were generated using \texttt{sim-cache}.
(c) Shows the average branch mispredictions per key 
incurred by our radix sort implementations while using bimodal predictors with 4096 table entries, the results for two-level adaptive predictors are extremely
similar. Finally, (d) shows the average number of branches per key executed by 
our  implementations. The results of (c) and (d) were generated using \texttt{sim-bpred}.}
\label{radixsort_results_pictures}
\end{figure}

Our radix sort implementation requires fewer processor cycles per key than even our base quicksort 
implementation as Figure \ref{radixsort_results_pictures}(a) shows. The number of level 2 cache misses is only slightly higher than we measured for
quicksort while the array fits inside the cache, presumably because radix sort is out-of-place.

Figure \ref{radixsort_results_pictures}(b) shows the level 2 cache misses for our basic and memory-tuned implementations.
We note that, when the array no longer fits inside the cache the memory-tuned implementation has approximately half the number of misses per key
(1.25) than the basic version (2.5). When the data no longer fits inside the cache there should be about one miss for each key per pass per cache line.
Each cache line can fit 8 keys, and thus we expect about 2 misses per key since there are 16 passes in total. The 2.5 misses per key which was measured
likely exceeds this because of the irregular fashion of the array indexing in each pass. Similary, in memory-tuned radix sort we expect about 1.125
misses per key, because there are 9 passes, again the measured 1.25 misses per key presumably exceeds this because
of the irregular array indexing.

The large advantage of radix sort is its almost complete lack of branch mispredictions however. Figure \ref{radixsort_results_pictures}(c) shows the tiny
number of branch mispredictions incurred by radix sort. It also executes a similarly tiny number of branch instructions as 
Figure \ref{radixsort_results_pictures}(d) shows.

\section{Related work} 
\label{RelatedWork}
% The IBM 7030 "Stretch" supercomputer was the first general-purpose
% pipelined architecture \cite{Bloch59}. Released in 1959 it had a
% four-stage pipeline and many other architectural innovations that are
% now commonplace in desktop and even embedded computers.

We present the first large-scale systematic study of branch prediction
and sorting. However, researchers have been aware for many years that
the branches in comparison-based sorts cause problems on pipelined
architectures. As early as 1972 Knuth commented on the
``ever-increasing number of `pipeline' or `number crunching' computers
that have appeared in recent years'' whose ``efficiency deteriorates
noticeably in the presence of conditional branch instructions unless
the branch almost always goes the same way''. He also notes that ``radix
sorting is usually more efficient than any other known method for
internal sorting on such machines'' (the same comment appears in \cite{KnuthVol3_98}). Most likely, Knuth
was referring to the IBM 7030 ``Stretch'' computer (released in 1961
with a four-stage pipeline), Seymour Cray's Freon-cooled CDC 6600
supercomputer (released 1964), or other early supercomputers which
used pipelining in the execution of general-purpose instructions.

A number of other authors have considered optimized implementations of
sorting algorithms for more recent single-processor machines. Nyberg
\textit{et al} \citeyear{Nyberg+94} present a quicksort-based algorithm for RISC
architectures. They found that the main limit on
performance was cache behaviour, and did not mention branch prediction
even once. Bentley and McIlroy \citeyear{Bentley+93} implemented a highly-tuned quicksort,
to be used in the \texttt{qsort} function of the standard C library. Again, they did not consider branch prediction,
perhaps because the first desktop processors with dynamic branch
predictors (such as the DEC Alpha 21064, MIPS R8000 and Intel Pentium)
were only just appearing around that time.

Agarwal \citeyear{Agarwal96} developed an optimized algorithm for sorting random records on
an IBM RS/6000 RISC machine. His is the first
published algorithm that we are aware of that deliberately attempts to
improve performance by eliminating branch mispredictions. First, it
uses the seven higher order bits of the key to divide the data into
128 buckets. Each of these buckets is radix-sorted. However, before
the radix sort is reaches the lowest order bits, the probability of
any adjacent pair of keys being out of order is extremely low, because
the problem states that the data is guaranteed to be random. In the
final stage, the algorithm checks that large runs of keys are
correctly ordered using subtraction and bitwise operations, and only
sorts a section if some key is found to be out of order. Although
Agarwal's techniques are interesting, they are very dependent on the
randomness of the data; it is easy to construct commmon cases that
would result in very poor performance. Nonetheless, the work is
interesting because it shows the effectiveness of radix sorting
techniques on pipelined architectures.

A recent trend in optimizing algorithms for FFT, linear algebra, and
digital signal processing has been to automatically generate and
evaluate very large numbers of variants of the algorithm.  A similar
approch is used by Li \textit{et al} \citeyear{Li+05} to automatically generate sorting
algorithms that are tuned to the architecture of the target machine
and the characteristics of the input data. They identify six sorting
primitives, which can be combined to construct a sorting algorithm.  A
large number of combinations are tested, using a genetic algorithms
search to find the most efficient. The resulting implementations are,
in many cases, significantly faster than commercial sorting libraries.
Although they do not consider branch prediction explicitly, they find
that the implementations that perform best usually use radix sort for
a large part of the sorting process.

Sanders and Winkel \citeyear{Sanders+04} investigate the use of \emph{predicated}
instructions on Intel Itanium 2 processors. In
addition to its normal inputs, a predicated instruction takes a
predicate register input. If the value in the predicate register is
true, the instruction takes effect; otherwise it does not. Predication
is normally implemented by the instruction being executed regardless
of whether the predicate is true. However, its result is only
\textit{written back} to a register or memory if the predicate is
true. Although predicated instructions use execution resources
regardless of whether they are allowed to write back, they allow
conditional branches to be eliminated.  Sanders and Winkel show how
the partition step in quicksort can be rewritten to use predicated
instructions on an Itanium 2, a highly instruction-level parallel
machine. Itanium 2 provides large numbers of parallel functional
units and registers which make such trade-offs worthwhile.

In  
preliminary experiments, we found that Sanders and Winkel's work is difficult 
to reproduce on a Pentium 4 based machine. The only predicated instruction
supported on the Pentium 4 is the conditional move instruction. We found that
rewriting
quicksort's loops so that they are suitable for predication
reduces their efficiency and also increases the number of simultaneously
live variables. Since the Pentium 4 has only eight general
purpose registers the number of simultaneously live variables exceeded
the number of available registers, reducing efficiency. In addition,
we also discovered that on the Pentium 4 conditional moves are
surprisingly expensive.

Mudge \textit{et al} \citeyear{Mudge+96} examine the behaviour of the $i$ and $j$
comparison branches in randomized quicksort, as an example to demonstrate a
general method for estimating the limit on the predictability of
branches in a program. They argue that the optimal
predictor for these branches would 
keep a running count of the total
number of array elements examined so far that are
greater than the pivot. If the majority of the elements examined so far
are greater than the pivot, the next element is predicted as greater
than the pivot, and vice versa. They estimate that this approach will give a
prediction accuracy of 75\%, which is close to our
result for these branches, using a bimodal predictor.
This shows that a bimodal predictor is close to optimal
for randomized quicksort's branches, since $P_{avg}$ of Section \ref{quicksort_branch_results_text}
is about 0.71.

Brodal \textit{et al} \citeyear{Brodal+05} examine the adaptiveness of quicksort.
They find that although randomized quicksort is not adaptive (i.e. its asymptotic analysis does not 
improve with respect to some measure of presortedness), its performance
is better on data with a low number of inversions. They justify this by showing 
that the expected number of element swaps performed by randomized quicksort falls as
the number of inversions does. The number of branch mispredictions
is roughly twice the number of element swaps, because two branch mispredictions
tend to occur when the two while loops of the partition code exit (see Figure \ref{partition_code}).
Furthermore they note that the branch mispredictions are the dominant part of the running time of
randomized quicksort. This provides an empirical relation between the presortedness of the input, 
randomized quicksort's performance, and the number of branch mispredictions.

In closely related work Brodal and Moruz \citeyear{BrodalMoruz05} provide a lower bound on the
number of branch mispredictions a comparison-based sorting algorithm must incur based on the
number of comparisons it performs. They also show that a mergesort variation can achieve this lower
bound based on the number of comparisons it performs, although they do not provide experimental
performance characteristics. They also extend their results to adaptive sorting algorithms

In recent work conducted independently of our own, Kaligosi and Sanders \citeyear{Kaligosi+06}
investigate the choice of pivot on the performance of quicksort. They examine $\alpha$-skewed
pivoting, that is, where the pivot has rank $\alpha n$ when sorting $n$ keys. They examine the performance
of quicksort when the pivot is determined randomly, from the median-of-3, or is the true median
(i.e. a 1/2-skewed pivot). They show that using a 1/10-skewed pivot gives a performance increase 
over all the aforementioned pivot choices. This is despite the increase in instruction count that a 
deliberately skewed pivot causes.

\section{Conclusion}

This paper has presented the first large scale study on the characteristics of branch instructions in 
all of the most common sorting algorithms. We noted that except where the sorting algorithms involve reasonably complicated
flow control (i.e. shellsort, multi-mergesort, multi-quicksort), bimodal predictors preform just as well as two-level adaptive predictors. Indeed we
noted that for shaker sort a bimodal predictor out-performs a two-level adaptive predictor.

Of the elementary sorting algorithms we noted that insertion sort,
causing just a single misprediction per key, has the best branch prediction characteristics. On the other hand,
bubble and shaker sort operate in a manner which causes surprisingly unpredictable branches.

Sorting (and indeed, searching) algorithms have been designed to minimize the number of comparisons necessary 
in the worst case. However, minimizing the number of comparisons makes each comparison less predictable. When a branch is
mispredicted, there is often a large penalty on modern processors. 
There is therefore a tension between reducing the number of executed comparisons and the penalty associated with 
increasing the chance of branch mispredictions. This tension is especially evident in the choice of pivot for quicksort.

Since cache optimizations are now recognized as very important, we 
have also noted the affect that these cache optimizations have on the branch prediction behaviour of sorting algorithms. In particular we
provided performance results for cache optimized versions of quicksort, mergesort, heapsort and radix sort.

Our study provides concrete results of the branch prediction characteristics for a large number
of sorting algorithms, accompanied with the relevant analysis.  These branch prediction results have been obtained for realistic dynamic predictors, using
a variety of configurations of bimodal and two-level adaptive predictors.

\appendix

\section{Appendix: Steady State Predictability}
\label{appendix_SSP}

In Sections \ref{bubble_sort}, \ref{heapsort_branch_results_text} and \ref{quicksort_branch_results_text}
we made use of the steady state predictability function $CP$. When a branch instruction is executed 
repeatedly with a known probability $p$ of being taken, $CP(p)$ gives a good estimate of the probability
that it will be correctly predicted.

The bimodal predictor associated with a branch instruction can be thought
of as a Markov chain. Note that we are thus assuming that successive branches
are independent. If the branch has probability $p$ of being taken then 
the transition matrix of the associated Markov chain is

\[
M = 
\left[
\begin{array}{llll}
p   & p   & 0   & 0 \\
1-p & 0   & p   & 0 \\
0   & 1-p & 0   & p \\
0   & 0   & 1-p & 1-p \\
\end{array}
\right]
\]

\noindent
The steady state probabilities are given by $s$ such that $Ms = s$.
Since the steady state probabilities sum to one this can be reduced to solving
a $3 \times 3$ linear system. The resultant probabilities are given by

\[
\left[ 
\begin{array}{c}
P(ST) \\
P(T) \\
P(NT) \\
P(SNT) \\
\end{array}
\right]
=
\frac{1}{2p^2 - 2p + 1}
\left[
\begin{array}{c}
p^3 \\
p^2(1 - p) \\
p(1 - p)^2 \\
(1 - p)^3 \\
\end{array}
\right]
\]

\noindent
Here $P(ST)$, $P(T)$, $P(NT)$ and $P(SNT)$ denote the respective probabilities of the predictor states strongly taken, taken, not-taken
and strongly not-taken described in Section \ref{branch_prediction}. A branch is correctly predicted with probability $p$ if we are 
in the strongly taken or taken state. Similarly, a branch is correctly predicted with probability $1 - p$ if we are in the strongly not-taken
or not-taken state. Thus,

\[
CP(p) = p\left[P(ST) + P(T)\right] + (1 - p)\left[P(SNT) + P(NT)\right]
\]

\noindent
This equation simplifies to Equation \ref{CP_equation} of Section \ref{bubble_sort}.
\bibliographystyle{acmtrans}
\bibliography{paper}
\begin{received}
\end{received}
\end{document}